### 2019-05-27 article/jdmdh-example.tex
Started article

\documentclass{jdmdh}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{pgfplots}

\title{Evaluating Deep Learning Methods for Tokenization of Scripta Continua in Old French and Latin}
\author[1]{Thibault Clérice}
\affil[1]{École nationale des Chartes, France} 
\affil[2]{Université Lyon 3, France}

\corrauthor{Thibault Clérice}{thibault.clerice@chartes.psl.eu}

\begin{document}

\maketitle

\abstract{Tokenization of modern and old Western European languages seems to be fairly simple as it stands on the presence mostly of markers such as spaces and punctuation. Although, when dealing with old sources like manuscript written in \textit{scripta continua}, (1) such markers are mostly absent, (2) spelling variation and rich morphologies makes dictionary based approaches difficult. We show that a convolutional encoder of characters followed by linear categorization can be used to tokenize such inputs. Additionaly, we release a software with a rather simple interface.}

\keywords{convolutional network; scripta continua; tokenization; Latin; Old French}

\section{Introduction}

\strut
\vspace{-4ex}

Sed eu tempor ipsum, vel cursus arcu. Maecenas non dignissim nunc, ac ornare tortor. Aenean
pretium arcu metus, id pulvinar enim tempus nec. Mauris faucibus mollis sodales. Sed
porttitor sed metus vitae vestibulum. Quisque a vehicula nunc. Aenean fringilla condimentum
diam, ac gravida quam. Integer ultrices feugiat enim nec tempus. Vestibulum ornare in magna
ultrices dapibus. Nulla facilisi.

\section{Title}

\subsection{Subtitle}
Pellentesque dignissim ultrices fringilla. Vivamus eu luctus ante, vel bibendum magna.
Curabitur elit purus, tincidunt non dui vitae, elementum bibendum neque. Curabitur
ullamcorper sit amet justo at hendrerit. Fusce ut arcu imperdiet nibh mollis tempus a aliquet
tellus. Quisque pharetra cursus nisi, vel lobortis ante consectetur et. Vivamus sed congue
neque. Proin pellentesque risus nec dui consequat rutrum. Vestibulum nunc diam, placerat
quis auctor vel, faucibus non justo. Etiam dictum purus neque. Phasellus imperdiet mauris
ligula, eu laoreet nisi elementum ut. Sed sed porta massa. Aenean faucibus risus ultrices
ornare porta. Quisque faucibus ante a tincidunt vestibulum. Lorem ipsum dolor sit amet,
consectetur adipiscing elit.

\subsubsection{Sub-subtitle}
Suspendisse vel dui nec felis molestie tincidunt. Vestibulum rutrum ligula lacus, ac molestie
nulla fermentum ornare. Nulla non nunc euismod, porta lacus vestibulum, malesuada massa.
Curabitur massa eros, rutrum sed lectus sed, volutpat semper metus. Mauris hendrerit aliquam
commodo.  Vivamus fermentum tempus pellentesque. Maecenas a hendrerit urna. In elit
ipsum, ultrices non dolor in, pulvinar porttitor lacus. Nunc euismod nibh quis odio
condimentum, a feugiat massa rutrum. Nulla erat erat, adipiscing vitae lectus id, consectetur
fermentum elit. Nunc eu est eu neque dapibus semper. Nam commodo urna dapibus, tincidunt
turpis a, cursus sem. Vivamus venenatis adipiscing mollis. Cras fringilla sodales lobortis.
Aliquam aliquet felis id est cursus auctor. Duis sodales tellus vulputate lectus egestas
volutpat.

\section{Tables and Figures}

\subsection{Table}
see Table~\ref{tab:example}.

\begin{table}
  \newcolumntype{+}{>{\global\let\currentrowstyle\relax}}
  \newcolumntype{^}{>{\currentrowstyle}}
  \newcommand{\rowstyle}[1]{\gdef\currentrowstyle{#1}%
    #1\ignorespaces
  }

  \centering
  \begin{tabular}{+>{\bfseries}l^c^c^c^c}
    \hline
    \rowstyle{\bfseries}
    & Sepal.Length & Sepal.Width & Petal.Length & Petal.Width\\
    Setosa & 5.006 & 3.428 & 1.462 & 0.246\\
    Versicolor & 5.936 & 2.77  & 4.26  & 1.326\\
    Verginica & 6.588 & 2.974 & 5.552 & 2.026\\
    \hline
  \end{tabular}

  \caption{Morbi malesuada diam at magna condimentum.}
  \label{tab:example}
\end{table}

\subsection{Figure}

Use subfigures to group similar images into one figure. Make pictures with a good resolution, as possible closed to 300 dpi or use vector graphics as for example provided by the TikZ package. Ensure that all legends are readable and in English. See example in Figure~\ref{fig:example} (taken from \url{http://www.texample.net/tikz/examples/pgfplots/}).

\begin{figure}
  \centering
  \begin{tikzpicture}
    \begin{axis}[
        xlabel=$x$,
        ylabel=$y$]
    \addplot[smooth,mark=*,blue] plot coordinates {
        (0,2)
        (2,3)
        (3,1)
    };
    \addlegendentry{Case 1}

    \addplot[smooth,color=red,mark=x]
        plot coordinates {
            (0,0)
            (1,1)
            (2,1)
            (3,2)
        };
    \addlegendentry{Case 2}
    \end{axis}
    \end{tikzpicture}

  \caption{Figure by Christian Feuers\"anger; Source: Pgfplots manual}
  \label{fig:example}
\end{figure}

\section{Definitions, Algorithms, and Formulas}

\subsection{Definitions}
You can define your own environments for formulas with the methods of the \texttt{amsthm} package. Use the theoremstyle \texttt{jdmdh} for a consistent formatting.

\begin{definition}[alpha]
Curabitur ullamcorper sit amet justo at hendrerit.
\end{definition}

Hence we set the following definition:

\begin{definition}
Etiam sed nulla viverra, ultrices ligula ac, consectetur libero.
\end{definition}

If you used a list, you need to write a bullet list:
\begin{itemize}
  \item Nunc id justo scelerisque.
  \item metus id enim iaculis tristique.
\end{itemize}

\subsection{Formulas}

Example of formula:
\begin{equation}
  Y=M.^tM-\beta.\langle M \rangle_l
\end{equation}

where $\langle M \rangle_l$ is a mean vector of a line from $M$. $\beta$ plays as a regulation factor to regulate the
rate of nearest neighbours, in fact the number of nearest neighbours is not defined explicitly.

Another example of a formula:
\begin{equation}
  K * N_c = Cst \pm 0.001\%
\end{equation}

\subsection{Algorithms}

This is a short algorithm:
\begin{lstlisting}[numbers=none]
quicksort(A, i, k):
  if i < k:
    p := partition(A, i, k)
    quicksort(A, i, p - 1)
    quicksort(A, p + 1, k)
\end{lstlisting}

For longer algorithms, use a float environment as shown in Listing~\ref{lst:example}. For configurations to specific languages, see the reference manual of the \texttt{listings} package.

\begin{listing}
  \begin{lstlisting}
  partition(array, left, right)
     pivotIndex := choose-pivot(array, left, right)
     pivotValue := array[pivotIndex]
     swap array[pivotIndex] and array[right]
     storeIndex := left
     for i from left to right - 1
         if array[i] < pivotValue
             swap array[i] and array[storeIndex]
             storeIndex := storeIndex + 1
     swap array[storeIndex] and array[right]  // Move pivot to its final place
     return storeIndex
  \end{lstlisting}

  \caption{Partition function of quicksort algorithm.}
  \label{lst:example}
\end{listing}

\section{References and Citations}
From \citet{sinclair91corpus} we pick up a general definition ...\\
\citet{ounis00flidar} explain that ...\\
\citet{wood92artifacts} recommend ...

% The \nocite command here is only intended to show more examples of references; please only list works that you really reference in your article
\nocite{hentschel07acceptance,biodiversa,pubmed,antonymy02perspective,justeson01cooccurence,r-project}

\subsection{Discussion}
Nam id eros massa. Fusce luctus purus a augue ullamcorper, sit amet vehicula mauris
tristique. Suspendisse eget pulvinar odio, nec bibendum turpis. Nullam quis lectus porttitor,
ullamcorper nisi et, condimentum leo. Quisque sed orci fermentum, rutrum velit eget, ultricies
augue. Nunc porttitor consectetur tincidunt. Nulla tincidunt justo enim, vitae dignissim erat
mattis ut. Nulla.

\subsection{Conclusion}
Maecenas egestas metus id enim iaculis tristique. Etiam sed nulla viverra, ultrices ligula ac,
consectetur libero. Nullam vitae massa ac odio pharetra condimentum. Maecenas in
elementum libero, non gravida quam. Praesent adipiscing consectetur consectetur. Vivamus at
orci sed augue varius hendrerit. Donec neque metus, dignissim nec erat at, ultricies consequat
libero. Donec eget eleifend leo. Aliquam at nunc porta, mollis sapien eu, eleifend tortor. Nam
egestas, metus ac pellentesque feugiat, lectus purus ornare est, vitae cursus felis turpis sit amet
lacus. Donec consequat massa mi, ac suscipit arcu posuere et. Vivamus et semper risus. Sed ut
arcu quam.

\bibliographystyle{plainnat}
\bibliography{jdmdh-example}

\appendix\footnotesize

\section{Annex 1}
Pellentesque dignissim ultrices fringilla. Vivamus eu luctus ante, vel bibendum magna. Curabitur elit purus, tincidunt non dui
vitae, elementum bibendum neque. Curabitur ullamcorper sit amet justo at hendrerit. Fusce ut arcu imperdiet nibh mollis
tempus a aliquet tellus. Quisque pharetra cursus nisi, vel lobortis ante consectetur et. Vivamus sed congue neque. Proin
pellentesque risus nec dui consequat rutrum. Vestibulum nunc diam, placerat quis auctor vel, faucibus non justo. Etiam
dictum purus neque. Phasellus imperdiet mauris ligula, eu laoreet nisi elementum ut. Sed sed porta massa. Aenean faucibus
risus ultrices ornare porta. Quisque faucibus ante a tincidunt vestibulum. Lorem ipsum dolor sit amet, consectetur adipiscing
elit.

\section{Annex 2}
Cras tristique vel nisi at aliquet. Proin egestas erat sit amet velit lobortis imperdiet. Integer et arcu sapien. Etiam id blandit
sapien. Nam tempus lacus ac massa semper, vel laoreet turpis rutrum. Mauris eget nibh vitae justo porta imperdiet sed vel
ligula. In imperdiet, augue vel condimentum convallis, neque augue imperdiet neque, eget dapibus nunc mauris ultricies
tortor. Nam eget nunc egestas, blandit lectus non, aliquam nunc. Cras sed quam vitae arcu ornare lobortis. Ut ut lacus
hendrerit, convallis orci sit amet, commodo nunc. Pellentesque eget tincidunt tortor. Nunc ornare molestie mauris id vehicula.
Suspendisse pharetra tortor metus, sit amet fermentum tellus vehicula ut.

\end{document}

### 2019-05-27 boudams/trainer.py

@@ -172,7 +172,7 @@ class Trainer(object):
             mode="loss",
             debug: Callable[[Seq2SeqTokenizer], None] = None
     ):
-        if seed:
+        if _seed:
             random.seed(_seed)
             torch.manual_seed(_seed)
             torch.backends.cudnn.deterministic = True

### 2019-05-27 boudams/model/base.py
Removed old stuff from previous Seq2Seq attempt

@@ -35,6 +35,9 @@ class BaseSeq2SeqModel(nn.Module):
         logits = torch.argmax(out, 2)[1:].t()  # Remove SOS token
         return label_encoder.reverse_batch(logits)
 
+    def init_weights(self):
+        pass
+
     def gradient(
         self,
         src, src_len, trg=None,

### 2019-05-27 boudams/model/bidir.py

@@ -59,220 +59,9 @@ class Encoder(nn.Module):
 
         return outputs, hidden
 
-
-class Attention(nn.Module):
-    def __init__(self, enc_hid_dim, dec_hid_dim):
-        super().__init__()
-
-        self.enc_hid_dim = enc_hid_dim
-        self.dec_hid_dim = dec_hid_dim
-
-        self.attn = nn.Linear((self.enc_hid_dim * 2) + self.dec_hid_dim, self.dec_hid_dim)
-        self.v = nn.Parameter(torch.rand(self.dec_hid_dim))
-
-    def forward(self, hidden, encoder_outputs, mask):
-        # hidden = [batch size, dec hid dim]
-        # encoder_outputs = [src sent len, batch size, enc hid dim * 2]
-        #mask = [batch size, src sent len]
-
-        batch_size = encoder_outputs.shape[1]
-        src_len = encoder_outputs.shape[0]
-
-        # repeat encoder hidden state src_len times
-        hidden = hidden.unsqueeze(1).repeat(1, src_len, 1)
-
-        encoder_outputs = encoder_outputs.permute(1, 0, 2)
-
-        # hidden = [batch size, src sent len, dec hid dim]
-        # encoder_outputs = [batch size, src sent len, enc hid dim * 2]
-
-        energy = torch.tanh(self.attn(torch.cat((hidden, encoder_outputs), dim=2)))
-
-        # energy = [batch size, src sent len, dec hid dim]
-
-        energy = energy.permute(0, 2, 1)
-
-        # energy = [batch size, dec hid dim, src sent len]
-
-        # v = [dec hid dim]
-
-        v = self.v.repeat(batch_size, 1).unsqueeze(1)
-
-        # v = [batch size, 1, dec hid dim]
-
-        attention = torch.bmm(v, energy).squeeze(1)
-
-        # attention= [batch size, src len]
-
-        attention = attention.masked_fill(mask == 0, -1e10)
-
-        return F.softmax(attention, dim=1)
-
-
-class Decoder(nn.Module):
-    def __init__(self, output_dim, emb_dim, enc_hid_dim, dec_hid_dim, dropout, attention):
-        super().__init__()
-
-        self.output_dim = output_dim
-        self.emb_dim = emb_dim
-        self.enc_hid_dim = enc_hid_dim
-        self.dec_hid_dim = dec_hid_dim
-        self.dropout = dropout
[…]

### 2019-05-27 boudams/model/conv.py

@@ -100,266 +100,3 @@ class Encoder(nn.Module):
 
         # combined = [batch size, src sent len, emb dim]
         return conved, combined
-
-
-class Decoder(nn.Module):
-    def __init__(self, output_dim, emb_dim, hid_dim, n_layers, kernel_size, dropout, pad_idx,
-                 max_sentence_len: int = 100,
-                 device: str = "cpu"):
-        super().__init__()
-
-        self.output_dim = output_dim
-        self.emb_dim = emb_dim
-        self.hid_dim = hid_dim
-        self.kernel_size = kernel_size
-        self.dropout = dropout
-        self.pad_idx = pad_idx
-        self.device = device
-        self.max_sentence_len = max_sentence_len
-
-        self.scale = torch.sqrt(torch.FloatTensor([0.5])).to(device)
-
-        self.tok_embedding = nn.Embedding(output_dim, emb_dim)
-        # Here, in the original code, it was 100
-        #   But POS_EMBEDDING first dimension should be the biggest size available
-        self.pos_embedding = nn.Embedding(self.max_sentence_len, emb_dim)
-
-        self.emb2hid = nn.Linear(emb_dim, hid_dim)
-        self.hid2emb = nn.Linear(hid_dim, emb_dim)
-
-        self.attn_hid2emb = nn.Linear(hid_dim, emb_dim)
-        self.attn_emb2hid = nn.Linear(emb_dim, hid_dim)
-
-        self.out = nn.Linear(emb_dim, output_dim)
-
-        self.convs = nn.ModuleList([nn.Conv1d(hid_dim, 2 * hid_dim, kernel_size)
-                                    for _ in range(n_layers)])
-
-        self.dropout = nn.Dropout(dropout)
-
-    def calculate_attention(self, embedded, conved, encoder_conved, encoder_combined):
-        # embedded = [batch size, trg sent len, emb dim]
-        # conved = [batch size, hid dim, trg sent len]
-        # encoder_conved = encoder_combined = [batch size, src sent len, emb dim]
-
-        # permute and convert back to emb dim
-        conved_emb = self.attn_hid2emb(conved.permute(0, 2, 1))
-
-        # conved_emb = [batch size, trg sent len, emb dim]
-
-        combined = (embedded + conved_emb) * self.scale
-
-        # combined = [batch size, trg sent len, emb dim]
-        energy = torch.matmul(combined, encoder_conved.permute(0, 2, 1))
-
[…]

### 2019-05-27 boudams/model/gru.py

@@ -1,179 +1,0 @@
-import torch
-import torch.nn as nn
-import random
-
-from .base import BaseSeq2SeqModel
-
-
-class Encoder(nn.Module):
-    def __init__(self, input_dim, emb_dim, hid_dim, dropout):
-        super().__init__()
-
-        self.input_dim = input_dim
-        self.emb_dim = emb_dim
-        self.hid_dim = hid_dim
-        self.dropout = dropout
-
-        self.embedding = nn.Embedding(input_dim, emb_dim)  # no dropout as only one layer!
-
-        self.rnn = nn.GRU(emb_dim, hid_dim)
-
-        self.dropout = nn.Dropout(dropout)
-
-    def forward(self, src, src_len):
-        # src = [src sent len, batch size]
-
-        embedded = self.dropout(self.embedding(src))
-
-        # embedded = [src sent len, batch size, emb dim]
-
-        packed_embedded = nn.utils.rnn.pack_padded_sequence(embedded, src_len)
-
-        packed_outputs, hidden = self.rnn(packed_embedded)  # no cell state!
-
-        outputs, _ = nn.utils.rnn.pad_packed_sequence(packed_outputs)
-
-        # outputs = [src sent len, batch size, hid dim * n directions]
-        # hidden = [n layers * n directions, batch size, hid dim]
-
-        # outputs are always from the top hidden layer
-
-        return hidden
-
-
-class Decoder(nn.Module):
-    def __init__(self, output_dim, emb_dim, hid_dim, dropout):
-        super().__init__()
-
-        self.emb_dim = emb_dim
-        self.hid_dim = hid_dim
-        self.output_dim = output_dim
-        self.dropout = dropout
-
-        self.embedding = nn.Embedding(output_dim, emb_dim)
-        self.rnn = nn.GRU(emb_dim + hid_dim, hid_dim)
-        self.out = nn.Linear(emb_dim + hid_dim * 2, output_dim)
-        self.dropout = nn.Dropout(dropout)
-
-    def forward(self, input, hidden, context):
-        # input = [batch size]
-        # hidden = [n layers * n directions, batch size, hid dim]
-        # context = [n layers * n directions, batch size, hid dim]
-
-        # n layers and n directions in the decoder will both always be 1, therefore:
-        # hidden = [1, batch size, hid dim]
-        # context = [1, batch size, hid dim]
-
-        input = input.unsqueeze(0)
-
-        # input = [1, batch size]
-
[…]

### 2019-05-27 boudams/model/linear.py

@@ -14,12 +14,18 @@ if TYPE_CHECKING:
 
 
 from .conv import Encoder as CNNEncoder
+from .lstm import Encoder as LSTMEncoder
+from .bidir import Encoder as BiGruEncoder
 
 
 class LinearEncoderCNN(CNNEncoder):
     def forward(self, src):
         out = super(LinearEncoderCNN, self).forward(src)
         return out
+
+
+class LinearLSTMEncoder(LSTMEncoder):
+    """ Linear version of the LSTMEncoder """
 
 
 class LinearDecoder(nn.Module):
@@ -82,12 +88,20 @@ class LinearSeq2Seq(BaseSeq2SeqModel):
         # calculate z^u (encoder_conved) and e (encoder_combined)
         # encoder_conved is output from final encoder conv. block
         # encoder_combined is encoder_conved plus (elementwise) src embedding plus positional embeddings
-        if self.encoder is not None:
+        if isinstance(self.encoder, LinearEncoderCNN):
             _, second_step = self.encoder(src)
+        elif isinstance(self.encoder, LinearLSTMEncoder):
+            second_step, hidden, cell = self.encoder(src.t(), src_len)
+            # -> tensor(sentence size, batch size, hid dim * n directions)
+            second_step = second_step.transpose(1, 0)
+            # -> tensor(batch size, sentence size, hid dim * n directions)
+        elif isinstance(self.encoder, BiGruEncoder):
+            second_step, hidden, cell = self.encoder(src.t(), src_len)
+            # -> tensor(sentence size, batch size, hid dim * n directions)
+            second_step = second_step.transpose(1, 0)
+            # -> tensor(batch size, sentence size, hid dim * n directions)
         else:
-            second_step = src
-        # encoder_conved = [batch size, src sent len, emb dim]
-        # encoder_combined = [batch size, src sent len, emb dim]
+            raise AttributeError("The encoder is not recognized.")
 
         output = self.decoder(second_step)
 
@@ -104,9 +118,11 @@ class LinearSeq2Seq(BaseSeq2SeqModel):
         """
         out = self(src, src_len, None, teacher_forcing_ratio=0)
         logits = torch.argmax(out, 2)
-        return label_encoder.reverse_batch(logits,
-                                           masked=override_src or src,
-                                           ignore=(self.pad_idx, self.eos_idx, self.sos_idx))
+        return label_encoder.reverse_batch(
+            logits,
[…]

### 2019-05-27 boudams/model/lstm.py

@@ -29,160 +29,14 @@ class Encoder(nn.Module):
 
         # embedded = [src sent len, batch size, emb dim]
 
-        packed_embedded = nn.utils.rnn.pack_padded_sequence(embedded, src_len)
-
         # packed_outputs = [src sent len, batch size, hid dim * n directions]
         # hidden = [n layers * n directions, batch size, hid dim]
         # cell = [n layers * n directions, batch size, hid dim]
-        packed_outputs, (hidden, cell) = self.rnn(packed_embedded)
-        #print(packed_outputs)
-        #print("--", flush=True)
-        #print(hidden)
-        #print("--", flush=True)
-        #print(cell)
-        #print("--", flush=True)
-        # packed_outputs is a packed sequence containing all hidden states
-        # hidden is now from the final non-padded element in the batch
+        output, (hidden, cell) = self.rnn(embedded)
 
-        outputs, _ = nn.utils.rnn.pad_packed_sequence(packed_outputs)
+        return output, hidden, cell
 
-        # outputs = [src sent len, batch size, hid dim * n directions]
-        # hidden = [n layers * n directions, batch size, hid dim]
-        # cell = [n layers * n directions, batch size, hid dim]
-
-        # outputs are always from the top hidden layer
-
-        return outputs, hidden, cell
-
-
-class Decoder(nn.Module):
-    def __init__(self, output_dim, emb_dim, hid_dim, n_layers, dropout):
-        super().__init__()
-
-        self.emb_dim = emb_dim
-        self.hid_dim = hid_dim
-        self.output_dim = output_dim
-        self.n_layers = n_layers
-        self.dropout = dropout
-
-        self.embedding = nn.Embedding(output_dim, emb_dim)
-
-        self.rnn = nn.LSTM(emb_dim, hid_dim, n_layers, dropout=dropout)
-
-        self.out = nn.Linear(hid_dim, output_dim)
-
-        self.dropout = nn.Dropout(dropout)
-
-    def forward(self, input, hidden, cell):
-        # input = [batch size]
-        # hidden = [n layers * n directions, batch size, hid dim]
-        # cell = [n layers * n directions, batch size, hid dim]
-
-        # n directions in the decoder will both always be 1, therefore:
-        # hidden = [n layers, batch size, hid dim]
-        # context = [n layers, batch size, hid dim]
-
-        input = input.unsqueeze(0)
-
-        # input = [1, batch size]
-
-        embedded = self.dropout(self.embedding(input))
-
-        # embedded = [1, batch size, emb dim]
-
[…]

### 2019-05-27 boudams/tagger.py

@@ -8,8 +8,8 @@ import logging
 import re
 from typing import List, Tuple
 
-from .model import gru, lstm, bidir, conv, linear
-from . import utils
+from boudams.model import gru, lstm, bidir, conv, linear
+from boudams import utils
 
 from .encoder import LabelEncoder, DatasetIterator
 
@@ -84,16 +84,25 @@ class Seq2SeqTokenizer:
         }
 
         if self.system.startswith("linear"):
-            self.enc: linear.CNNEncoder = linear.CNNEncoder(
-                    self.vocabulary_dimension, emb_dim=self.emb_enc_dim,
-                    n_layers=self.enc_n_layers, hid_dim=self.enc_hid_dim,
-                    dropout=self.enc_dropout,
-                    device=self.device,
-                    kernel_size=self.enc_kernel_size,
-                    max_sentence_len=self.out_max_sentence_length
-                )
+            if self.system.endswith("-lstm"):
+                self.enc: linear.LSTMEncoder = linear.LinearLSTMEncoder(
+                        self.vocabulary_dimension, emb_dim=self.emb_enc_dim,
+                        n_layers=self.enc_n_layers, hid_dim=self.enc_hid_dim,
+                        dropout=self.enc_dropout
+                    )
+                in_features = self.enc_hid_dim
+            else:
+                self.enc: linear.CNNEncoder = linear.LinearEncoderCNN(
+                        self.vocabulary_dimension, emb_dim=self.emb_enc_dim,
+                        n_layers=self.enc_n_layers, hid_dim=self.enc_hid_dim,
+                        dropout=self.enc_dropout,
+                        device=self.device,
+                        kernel_size=self.enc_kernel_size,
+                        max_sentence_len=self.out_max_sentence_length
+                    )
+                in_features = self.emb_enc_dim
             self.dec: linear.LinearDecoder = linear.LinearDecoder(
-                enc_dim=self.emb_enc_dim, out_dim=len(self.vocabulary.mtoi)
+                enc_dim=in_features, out_dim=len(self.vocabulary.mtoi)
             )
             self.model: linear.LinearSeq2Seq = linear.LinearSeq2Seq(
                 self.enc, self.dec, **seq2seq_shared_params

### 2019-05-28 boudams/cli.py
Adding dataset and clean-up of remaining items from Seq2Seq

@@ -187,5 +187,40 @@ def tag(model, filename, device="cpu", batch_size=64):
         # print("--- File " + file.name + " has been tokenized")
 
 
+@cli.command("graph")
+@click.argument("model", type=click.Path(exists=True, file_okay=True, dir_okay=False))
+@click.argument("output", type=click.Path(file_okay=True, dir_okay=False))
+@click.option("--format", default="png", type=click.Choice("png", "pdf"))
+def graph(model, output, format):
+    """ Draw the graph representation of a given model """
+    try:
+        import hiddenlayer as hl
+        import hiddenlayer.pytorch_builder as torch_builder
+    except ImportError:
+        print("You need to install hiddenlayer (pip install hiddenlayer) for this command.")
+        return
+    import torch
+
+    print("Loading the model.")
+    model = Seq2SeqTokenizer.load(model, device="cpu")
+    print("Model loaded.")
+
+    tensor = torch.ones((model.out_max_sentence_length,), dtype=torch.float64)
+
+    # Build HiddenLayer graph
+    g = hl.Graph()
+    hl_graph = torch_builder.import_graph(
+        g,
+        model.model,
+        args=(
+            torch.zeros([64, model.out_max_sentence_length], dtype=torch.long),
+            tensor.new_full((64, ), model.out_max_sentence_length)
+        ))
+
+    # Use a different color theme
+    hl_graph.theme = hl.graph.THEMES["blue"].copy()  # Two options: basic and blue
+    hl_graph.save(output, format=format)
+
+
 if __name__ == "__main__":
     cli()

### 2019-05-28 boudams/model/conv.py

@@ -66,6 +66,7 @@ class Encoder(nn.Module):
 
         # conv_input = [batch size, src sent len, hid dim]
 
+
         # permute for convolutional layer
         conv_input = conv_input.permute(0, 2, 1)
 
@@ -87,7 +88,7 @@ class Encoder(nn.Module):
 
             # conved = [batch size, hid dim, src sent len]
 
-            # set conv_input to conved for next loop iteration
+            # set conv_input to conved for next lo`op iteration
             conv_input = conved
 
         # permute and convert back to emb dim

### 2019-05-28 boudams/model/linear.py

@@ -81,7 +81,7 @@ class LinearSeq2Seq(BaseSeq2SeqModel):
         nll_weight[pad_idx] = 0.
         self.register_buffer('nll_weight', nll_weight)
 
-    def forward(self, src, src_len, trg, **kwargs):
+    def forward(self, src, src_len, trg=None, **kwargs):
         # src = [batch size, src sent len]
         # trg = [batch size, trg sent len]
 

### 2019-05-28 boudams/tagger.py

@@ -8,7 +8,7 @@ import logging
 import re
 from typing import List, Tuple
 
-from boudams.model import gru, lstm, bidir, conv, linear
+from boudams.model import lstm, bidir, conv, linear
 from boudams import utils
 
 from .encoder import LabelEncoder, DatasetIterator
@@ -25,11 +25,11 @@ class Seq2SeqTokenizer:
             self,
             vocabulary: LabelEncoder,
             hidden_size: int = 256,
-            enc_n_layers: int = 10, dec_n_layers: int = 10,
-            emb_enc_dim: int = 256, emb_dec_dim: int = 256,
-            enc_hid_dim: int = None, dec_hid_dim: int = None,
-            enc_dropout: float = 0.5, dec_dropout: float = 0.5,
-            enc_kernel_size: int = 3, dec_kernel_size: int = 3,
+            enc_n_layers: int = 10,
+            emb_enc_dim: int = 256,
+            enc_hid_dim: int = None,
+            enc_dropout: float = 0.5,
+            enc_kernel_size: int = 3,
             out_max_sentence_length: int = 150,
             device: str = DEVICE, system: str = "bi-gru",
             **kwargs # RetroCompat
@@ -52,18 +52,13 @@ class Seq2SeqTokenizer:
         self.device: str = device
         self.enc_hid_dim = self.dec_hid_dim = self.hidden_size = hidden_size
 
-        if enc_hid_dim and dec_hid_dim:
+        if enc_hid_dim:
             self.enc_hid_dim: int = enc_hid_dim
-            self.dec_hid_dim: int = dec_hid_dim
 
         self.emb_enc_dim: int = emb_enc_dim
-        self.emb_dec_dim: int = emb_dec_dim
         self.enc_dropout: float = enc_dropout
-        self.dec_dropout: float = dec_dropout
         self.enc_kernel_size: int = enc_kernel_size
-        self.dec_kernel_size: int = dec_kernel_size
         self.enc_n_layers: int = enc_n_layers
-        self.dec_n_layers: int = dec_n_layers
 
         self.out_max_sentence_length: int = out_max_sentence_length
         self.system: str = system
@@ -83,90 +78,31 @@ class Seq2SeqTokenizer:
             "out_max_sentence_length": self.out_max_sentence_length
         }
 
-        if self.system.startswith("linear"):
-            if self.system.endswith("-lstm"):
[…]

### 2019-05-28 setup.py

@@ -110,7 +110,7 @@ setup(
     },
     install_requires=REQUIRED,
     extras_require=EXTRAS,
-    include_package_data=True,
+    include_package_data=False,
     license='MIT',
     classifiers=[
         # Trove classifiers
@@ -119,8 +119,6 @@ setup(
         'Programming Language :: Python',
         'Programming Language :: Python :: 3',
         'Programming Language :: Python :: 3.6',
-        'Programming Language :: Python :: Implementation :: CPython',
-        'Programming Language :: Python :: Implementation :: PyPy',
         'Topic :: Text Processing :: Linguistic'
     ],
     # $ setup.py publish support.

### 2019-05-28 test_data/himanis/convert_LIRIS.py

@@ -1,0 +1,12 @@
+import os
+
+directory = os.path.dirname(__file__)
+lang = "lat"
+
+with open(os.path.join(directory, "txt", "LIRIS_mss_dates_w.txt")) as f:
+	for line_index, line in enumerate(f.readlines()):
+		if not line.strip():
+			continue
+			#
+		with open(os.path.join(directory, lang, "LIRIS-%s.txt"% line_index), "w") as out_io:
+			out_io.write(line.strip())

### 2019-05-28 test_data/himanis/convert_guerin.py

@@ -1,0 +1,13 @@
+import os
+
+directory = os.path.dirname(__file__)
+
+with open(os.path.join(directory, "txt", "guerin.txt")) as f:
+	for line_index, line in enumerate(f.readlines()):
+
+		if not line.strip():
+			continue
+			#
+		lang, text = line[:3], line[7:].strip()
+		with open(os.path.join(directory, lang, "guerin-%s.txt"% line_index), "w") as out_io:
+			out_io.write(text)

### 2019-05-28 boudams/cli.py
Fixed some issue with some of the dataset

@@ -113,56 +113,57 @@ def template(filename):
 
 
 @cli.command("train")
-@click.argument("config_file", type=click.File("r"))
+@click.argument("config_files", nargs=-1, type=click.File("r"))
 @click.option("--epochs", type=int, default=100, help="Number of epochs to run")
 @click.option("--batch_size", type=int, default=32, help="Size of batches")
 @click.option("--device", default="cpu", help="Device to use for the network (cuda, cpu, etc.)")
-def train(config_file, epochs, batch_size, device):
-    """ Train a model """
-    config = json.load(config_file)
+def train(config_files, epochs, batch_size, device):
+    """ Train one or more models according to [CONFIG_FILES] JSON configurations"""
+    for config_file in config_files:
+        config = json.load(config_file)
 
-    masked = config["model"].startswith("linear")
-    train_path, dev_path, test_path = config["datasets"]["train"],\
-                                      config["datasets"]["dev"],\
-                                      config["datasets"]["test"]
+        masked = config["model"].startswith("linear")
+        train_path, dev_path, test_path = config["datasets"]["train"],\
+                                          config["datasets"]["dev"],\
+                                          config["datasets"]["test"]
 
-    vocabulary = LabelEncoder(
-        maximum_length=config["max_sentence_size"],
-        masked=masked,
-        remove_diacriticals=config["label_encoder"].get("normalize", True),
-        lower=config["label_encoder"].get("lower", True)
-    )
-    vocabulary.build(train_path, dev_path, test_path, debug=True)
+        vocabulary = LabelEncoder(
+            maximum_length=config["max_sentence_size"],
+            masked=masked,
+            remove_diacriticals=config["label_encoder"].get("normalize", True),
+            lower=config["label_encoder"].get("lower", True)
+        )
+        vocabulary.build(train_path, dev_path, test_path, debug=True)
 
-    # Get the datasets
-    train_dataset: DatasetIterator = vocabulary.get_dataset(
-        train_path, randomized=config["datasets"].get("random", True))
-    dev_dataset: DatasetIterator = vocabulary.get_dataset(
-        dev_path, randomized=config["datasets"].get("random", True))
-    test_dataset: DatasetIterator = vocabulary.get_dataset(
-        test_path, randomized=config["datasets"].get("random", True))
+        # Get the datasets
+        train_dataset: DatasetIterator = vocabulary.get_dataset(
[…]

### 2019-05-28 boudams/dataset/plaintext.py

@@ -2,7 +2,7 @@ import glob
 import os
 import os.path
 import random
-import re
+import regex as re
 
 from typing import Iterable, Union
 

### 2019-05-28 boudams/encoder.py

@@ -7,6 +7,7 @@ import collections
 import random
 import json
 import unidecode
+import copy
 from operator import itemgetter
 
 DEVICE = torch.device("cuda" if torch.cuda.is_available() else "cpu")
@@ -17,7 +18,7 @@ DEFAULT_UNK_TOKEN = "<UNK>"
 DEFAULT_MASK_TOKEN = "x"
 
 
-GT_PAIR = collections.namedtuple("GT", ("x", "x_length", "y", "y_length"))
+GT_PAIR = collections.namedtuple("GT", ("x", "x_length", "y", "y_length", "line_index"))
 
 
 class DatasetIterator:
@@ -76,12 +77,15 @@ class DatasetIterator:
         """
         logging.info("DatasetIterator reading indexes of lines")
         with open(self.file, "r") as fio:
-            for line in fio.readlines():
+            for line_index, line in enumerate(fio.readlines()):
+                if not line.strip():
+                    continue
                 x, y = self._l_e.readunit(line.strip())
                 self.encoded.append(
                     GT_PAIR(
                         *self._l_e.inp_to_numerical(x),
-                        *self._l_e.gt_to_numerical(y)
+                        *self._l_e.gt_to_numerical(y),
+                        line_index
                     )
                 )
 
@@ -117,13 +121,17 @@ class DatasetIterator:
                     xs.append(gt_pair.x)
                     y_trues.append(gt_pair.y)
 
-                x_tensor, x_length, x_order = self._l_e.pad_and_tensorize(xs, padding=max_len_x, device=device)
-                y_tensor, y_length, y_order = self._l_e.pad_and_tensorize(
-                    y_trues,
-                    padding=max_len_y,
-                    device=device,
-                    reorder=x_order
-                )
+                try:
+                    x_tensor, x_length, x_order = self._l_e.pad_and_tensorize(xs, padding=max_len_x, device=device)
+                    y_tensor, y_length, y_order = self._l_e.pad_and_tensorize(
+                        y_trues,
+                        padding=max_len_y,
+                        device=device,
+                        reorder=x_order
+                    )
+                except ValueError:
+                    print([b.line_index for b in lines[n:n+self.batch_size]])
+                    raise
                 yield (
                     x_tensor,
                     x_length,
@@ -290,6 +298,7 @@ class LabelEncoder:
         lengths = []
         order = []
 
+        original_order = copy.deepcopy(sentences)
         # If GT order was computer, we get it from there
         if reorder:
             sequences = [sentences[index] for index in reorder]
[…]

### 2019-05-28 boudams/model/linear.py

@@ -19,9 +19,11 @@ from .bidir import Encoder as BiGruEncoder
 
 
 class LinearEncoderCNN(CNNEncoder):
-    def forward(self, src):
-        out = super(LinearEncoderCNN, self).forward(src)
-        return out
+    def forward(self, src, keep_pos=False):
+        o, p = super(LinearEncoderCNN, self).forward(src)
+        if keep_pos:
+            return p
+        return o
 
 
 class LinearLSTMEncoder(LSTMEncoder):
@@ -64,12 +66,14 @@ class LinearSeq2Seq(BaseSeq2SeqModel):
         encoder: CNNEncoder, decoder: LinearDecoder,
         device: str,
         pad_idx: int, sos_idx: int, eos_idx: int,
+        pos: bool = False,
         **kwargs
     ):
         super().__init__()
 
         self.encoder = encoder
         self.decoder: LinearDecoder = decoder
+        self.pos = pos
 
         self.pad_idx = pad_idx
         self.sos_idx = sos_idx
@@ -89,7 +93,7 @@ class LinearSeq2Seq(BaseSeq2SeqModel):
         # encoder_conved is output from final encoder conv. block
         # encoder_combined is encoder_conved plus (elementwise) src embedding plus positional embeddings
         if isinstance(self.encoder, LinearEncoderCNN):
-            _, second_step = self.encoder(src)
+            second_step = self.encoder(src, keep_pos=self.pos)
         elif isinstance(self.encoder, LinearLSTMEncoder):
             second_step, hidden, cell = self.encoder(src.t(), src_len)
             # -> tensor(sentence size, batch size, hid dim * n directions)

### 2019-05-28 boudams/tagger.py

@@ -100,7 +100,9 @@ class Seq2SeqTokenizer:
             enc_dim=in_features, out_dim=len(self.vocabulary.mtoi)
         )
         self.model: linear.LinearSeq2Seq = linear.LinearSeq2Seq(
-            self.enc, self.dec, **seq2seq_shared_params
+            self.enc, self.dec,
+            pos="nopos" not in self.system,
+            **seq2seq_shared_params
         ).to(device)
         self.init_weights = None
 

### 2019-05-28 generate.sh

@@ -1,0 +1,9 @@
+#!/usr/bin/env bash
+
+venv/bin/boudams dataset convert plain-text test_data/bfmmss/gt test_data/bfmmss/txt/*.txt
+venv/bin/boudams dataset convert plain-text test_data/geste/gt test_data/geste/src/*.txt
+venv/bin/boudams dataset convert plain-text test_data/bnf412/gt test_data/bnf412/txt/*.txt
+venv/bin/boudams dataset convert plain-text test_data/himanis/gt/fro test_data/himanis/fro/*.txt
+venv/bin/boudams dataset convert plain-text test_data/pinche/gt test_data/pinche/src/*.txt
+
+venv/bin/boudams dataset generate datasets/fro test_data/bfmmss/gt/*.txt test_data/geste/gt/*.txt test_data/bnf412/gt/*.txt test_data/himanis/gt/fro/*.txt test_data/pinche/gt/*.txt

### 2019-05-28 test_data/bugged.py

@@ -1,0 +1,14 @@
+from boudams.encoder import LabelEncoder
+import logging
+
+
+logger = logging.getLogger()
+logger.setLevel(logging.DEBUG)
+
+label = LabelEncoder(masked=True)
+label.build("bugged.txt", debug=True)
+
+dataset = label.get_dataset("bugged.txt")
+batch_generator = dataset.get_epoch(batch_size=32)()
+batch = next(batch_generator)
+print(batch)

### 2019-05-28 test_data/bugged.py
Removed zero width no break space

@@ -5,10 +5,10 @@ import logging
 logger = logging.getLogger()
 logger.setLevel(logging.DEBUG)
 
-label = LabelEncoder(masked=True)
+label = LabelEncoder(masked=True, lower=False, remove_diacriticals=True)
 label.build("bugged.txt", debug=True)
 
 dataset = label.get_dataset("bugged.txt")
-batch_generator = dataset.get_epoch(batch_size=32)()
+batch_generator = dataset.get_epoch(batch_size=64)()
 batch = next(batch_generator)
 print(batch)

### 2019-05-29 article/article.tex
Some changes here and there

\documentclass{jdmdh}
\usepackage[utf8]{inputenc}
\usepackage{array}
\usepackage{pgfplots}

\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French and Latin}
\author[1]{Thibault Clérice}
\affil[1]{École nationale des Chartes, France} 
\affil[2]{Université Lyon 3, France}

\corrauthor{Thibault Clérice}{thibault.clerice@chartes.psl.eu}

\begin{document}

\maketitle

\abstract{Tokenization of modern and old Western European languages seems to be fairly simple as it stands on the presence mostly of markers such as spaces and punctuation. Although, when dealing with old sources like manuscript written in \textit{scripta continua}, (1) such markers are mostly absent, (2) spelling variation and rich morphology makes dictionary based approaches difficult. We show that applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence can be used to tokenize such inputs. Additionally, we release a software with a rather simple interface for tokenizing one's corpus.}

\keywords{convolutional network; scripta continua; tokenization; Latin; Old French; word segmentation}

\section{Introduction}

% To Read : Stutzmann article.

Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando". \textit{Scripta continua} is a writing phenomenon where words would not be separated by spaces and it appears to have disappeared around the 8th century (see \citet{zanna1998lecture}). Although, spacing can be somewhat erratic in later centuries (\textit{cf.} Figure \ref{fig:4lines}). In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{4-lines-p0215.png}

  \caption{ 4 lines from fol.103rb Manuscript fr. 412, Bibliothèque nationale de France.  Red lines indicate word boundaries}
  \label{fig:4lines}
\end{figure}

We must stress in this study that the difficulty that we face is different for \textit{scripta continua} than the ones researchers face languages such as Chinese for which an already impressive amount of work has been done as it. Indeed, Chinese word segmentation has lately been driven by deep learning methods, specifically ones based on \textit{sequence to sequence translations}: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses BiDirectional GRU and CRF. Actually, meanwhile redacting this article and producing the code-base, \citet{huang2019realistic} took the same approach of encoding to linear classification to both word boundary (WB) and word content (WC) for Chinese word segmentation.

\section{Description and evaluation}

\subsection{Architecture}

\subsubsection{Encoding of input and decoding}

The model is based on traditional text input encoding where each character is transcoded to an index. Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content.

\begin{table}[]
\centering
\begin{tabular}{@{}ll@{}}
\hline
                       & \textbf{Sample}           \\  \hline
\textbf{Input  String} & \texttt{Ladamehaitees'enparti}     \\
\textbf{Mask   String} & \texttt{xSxxxSxxxxxSxxxSxxxxS}     \\
\textbf{Output String} & \texttt{La dame haitee s'en parti} \\ \hline
\end{tabular}
  \caption{Input, mask and human-readable output generated by the model. x are WC and S are WB}
  \label{lst:input_output_example}
\end{table}

For evaluation purposes, and to reduce the number of input classes, we propose both a lower-case normalization and a "reduction to the ASCII character set" feature (fr. \ref{fig:normalization}). On this point, a lot of issues were found with transliteration of MUFI characters that were part of the original datasets, as they are badly interpreted by the \texttt{unidecode} python package. Indeed, \texttt{unidecode} will simply remove characters it does not understand. An addendum to said package that would transcribe ligatures, abbreviations, etc. would probably be a good feature.

% Need a better system that this/
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{inputs.png}
  \caption{ Different possibilities of pre-processing. In the last, the first character (LATIN ABBREVIATION SIGN SMALL ET WITH STROKE) is lost by \texttt{unidecode}}
  \label{fig:normalization}
\end{figure}

\subsubsection{Model}

Each model we have used is composed of an Encoder, generally built around an embedding layer, and the same Linear Classifier. The encoders tested are:

\begin{itemize}
  \item LSTM encoder with hidden cell
  \item Convolutional (CNN) encoder with position embeddings 
  \item Convolutional (CNN) encoder without position embeddings % Currently not really because it's part of the network, but it should be for the release of the article
\end{itemize}

\subsection{Evaluation}

\subsubsection{Datasets}

Datasets are transcription from manuscripts with unresolved abbreviation coming from different projects

\begin{itemize}
    \item \textbf{Old French} based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and transcription from the TNAH master at the École Nationale des Chartes.
    \begin{itemize}
        \item 193,734 training examples;
        \item 23,581 validation examples;
        \item 25,512 test examples
    \end{itemize}
\end{itemize}

\subsubsection{Results}

\begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
          width=\linewidth, % Scale the plot to \linewidth
          grid=major, 
          grid style={dashed,gray!30},
          xlabel=Epoch, % Set the labels
          ylabel=Accuracy,
          legend style={at={(0.5,-0.2)},anchor=north},
          x tick label style={rotate=90,anchor=east}
        ]
        \addplot table[x=Epoch,y=CNN Train,col sep=comma] {accuracies.csv}; 
        \addplot table[x=Epoch,y=CNN No Pos,col sep=comma] {accuracies.csv}; 
        \addplot table[x=Epoch,y=CNN lower,col sep=comma] {accuracies.csv}; 
        \addplot table[x=Epoch,y=CNN norm,col sep=comma] {accuracies.csv}; 
        \legend{CNN Train,CNN without Position embedding, CNN with lowercase transform, CNN with ASCIIification}
      \end{axis}
    \end{tikzpicture}
    \caption{Training Accuracy over 50 epochs}
  \label{fig:accuracy}
  \end{center}
\end{figure}

% Include evaluation on test sets

\subsubsection{Example of outputs}

The following input was not seen during training, dev or test, and is part of a different corpus :

\begin{itemize} 
    \item \texttt{Truth :} Aies joie et leesce en ton cuer car tu auras une fille qui aura .i. fil qui sera de molt grant merite devant Dieu et de grant los entre les homes.
    \item \texttt{Input :} Aiesjoieetleesceentoncuercartuaurasunefillequiaura.i.filquiserademoltgrantmeritedevantDieuetdegrantlosentreleshomes.
    \item \texttt{CNN :} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN lower:} Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN without position:} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
    \item \texttt{CNN Normalize:} Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
\end{itemize}

\subsection{Discussion}

We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view

% "The purpose of the discussion is to interpret and describe the significance of your findings in light of what was already known about the research problem being investigated and to explain any new understanding or insights that emerged as a result of your study of the problem. The discussion will always connect to the introduction by way of the research questions or hypotheses you posed and the literature you reviewed, but the discussion does not simply repeat or rearrange the first parts of your paper; the discussion clearly explain how your study advanced the reader's understanding of the research problem from where you left them at the end of your review of prior research."

\subsection{Conclusion}

While 
% The conclusion is intended to help the reader understand why your research should matter to them after they have finished reading the paper. A conclusion is not merely a summary of the main topics covered or a re-statement of your research problem, but a synthesis of key points and, if applicable, where you recommend new areas for future research. For most college-level research papers, one or two well-developed paragraphs is sufficient for a conclusion, although in some cases, three or more paragraphs may be required.

\subsection{Acknowledgement}

Boudams has been made possible by two open-source repositories from which I learned and copied bits of implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}. This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.

\bibliographystyle{plainnat}
\bibliography{article}

\appendix\footnotesize

\section{Annex 1}
Pellentesque dignissim ultrices fringilla. Vivamus eu luctus ante, vel bibendum magna. Curabitur elit purus, tincidunt non dui
vitae, elementum bibendum neque. Curabitur ullamcorper sit amet justo at hendrerit. Fusce ut arcu imperdiet nibh mollis
tempus a aliquet tellus. Quisque pharetra cursus nisi, vel lobortis ante consectetur et. Vivamus sed congue neque. Proin
pellentesque risus nec dui consequat rutrum. Vestibulum nunc diam, placerat quis auctor vel, faucibus non justo. Etiam
dictum purus neque. Phasellus imperdiet mauris ligula, eu laoreet nisi elementum ut. Sed sed porta massa. Aenean faucibus
risus ultrices ornare porta. Quisque faucibus ante a tincidunt vestibulum. Lorem ipsum dolor sit amet, consectetur adipiscing
elit.

\section{Annex 2}
Cras tristique vel nisi at aliquet. Proin egestas erat sit amet velit lobortis imperdiet. Integer et arcu sapien. Etiam id blandit
sapien. Nam tempus lacus ac massa semper, vel laoreet turpis rutrum. Mauris eget nibh vitae justo porta imperdiet sed vel
ligula. In imperdiet, augue vel condimentum convallis, neque augue imperdiet neque, eget dapibus nunc mauris ultricies
tortor. Nam eget nunc egestas, blandit lectus non, aliquam nunc. Cras sed quam vitae arcu ornare lobortis. Ut ut lacus
hendrerit, convallis orci sit amet, commodo nunc. Pellentesque eget tincidunt tortor. Nunc ornare molestie mauris id vehicula.
Suspendisse pharetra tortor metus, sit amet fermentum tellus vehicula ut.

\end{document}

### 2019-05-29 boudams/cli.py

@@ -142,7 +142,7 @@ def train(config_files, epochs, batch_size, device):
             dev_path, randomized=config["datasets"].get("random", True))
         test_dataset: DatasetIterator = vocabulary.get_dataset(
             test_path, randomized=config["datasets"].get("random", True))
-
+        print("Training %s " % config_file.name)
         print("-- Dataset informations --")
         print("Number of training examples: {}".format(len(train_dataset)))
         print("Number of dev examples: {}".format(len(dev_dataset)))
@@ -187,6 +187,17 @@ def tag(model, filename, device="cpu", batch_size=64):
                 out_io.write(tokenized_string+" ")
         # print("--- File " + file.name + " has been tokenized")
 
+
+@cli.command("tag-check")
+@click.argument("config_model", nargs=-1, type=click.Path(exists=True, file_okay=True, dir_okay=False))
+@click.argument("content")
+def tag_check(config_model, content, device="cpu", batch_size=64):
+    """ Tag all [FILENAME] using [MODEL]"""
+    for model in config_model:
+        print("Loading the model.")
+        tokenizer = Seq2SeqTokenizer.load(model, device=device)
+        print("Model loaded.")
+        print(model + "\t" +" ".join(tokenizer.annotate_text(content, batch_size=batch_size)))
 
 @cli.command("graph")
 @click.argument("model", type=click.Path(exists=True, file_okay=True, dir_okay=False))

### 2019-05-29 boudams/encoder.py

@@ -6,9 +6,10 @@ import logging
 import collections
 import random
 import json
-import unidecode
 import copy
 from operator import itemgetter
+
+from boudams.utils import mufidecode
 
 DEVICE = torch.device("cuda" if torch.cuda.is_available() else "cpu")
 DEFAULT_INIT_TOKEN = "<SOS>"
@@ -259,7 +260,7 @@ class LabelEncoder:
             x = x.lower()
 
         if self.remove_diacriticals:
-            x = unidecode.unidecode(x)
+            x = mufidecode(x.replace("ꝯ", "9")).strip()
 
         x = x.split("\t")
 
@@ -272,7 +273,7 @@ class LabelEncoder:
             x = x.lower()
 
         if self.remove_diacriticals:
-            x = unidecode.unidecode(x)
+            x = mufidecode(x.replace("ꝯ", "9")).strip()
         return x
 
     def pad_and_tensorize(
@@ -330,7 +331,7 @@ class LabelEncoder:
                     self.mask_token_index if ngram[1] != " " else self.space_token_index
                     for ngram in zip(*[sentence[i:] for i in range(2)])
                     if ngram[0] != " "
-                ] + [self.mask_token_index] + eos
+                ] + [self.space_token_index] + eos
 
             return numericals, len(sentence) - sentence.count(" ") + obligatory_tokens
 

### 2019-05-29 boudams/utils.py

@@ -5,10 +5,52 @@ import uuid
 from contextlib import contextmanager
 import os
 import shutil
+import unidecode
+import warnings
 
-import matplotlib.pyplot as plt
-import matplotlib.ticker as ticker
-plt.switch_backend('agg')
+
+Cache = {}
+
+
+def mufidecode(string):
+    retval = []
+
+    for char in string:
+        codepoint = ord(char)
+
+        if codepoint < 0x80:  # Basic ASCII
+            retval.append(str(char))
+            continue
+
+        if 0xd800 <= codepoint <= 0xdfff:
+            warnings.warn("Surrogate character %r will be ignored. "
+                          "You might be using a narrow Python build." % (char,),
+                          RuntimeWarning, 2)
+
+        section = codepoint >> 8  # Chop off the last two hex digits
+        position = codepoint % 256  # Last two hex digits
+
+        try:
+            table = Cache[section]
+        except KeyError:
+            try:
+                mod = __import__('unidecode.x%03x' % (section), globals(), locals(), ['data'])
+            except ImportError:
+                Cache[section] = None
+                retval.append(char)
+                continue
+                
+            Cache[section] = table = mod.data
+
+        if table and len(table) > position:
+            retval.append(table[position])
+            continue
+
+        # If not found
+        retval.append(char)
+        continue
+
+    return ''.join(retval)
 
 
 def epoch_time(start_time, end_time):
@@ -17,14 +59,6 @@ def epoch_time(start_time, end_time):
     elapsed_secs = int(elapsed_time - (elapsed_mins * 60))
     return elapsed_mins, elapsed_secs
 
-
-def showPlot(points):
-    plt.figure()
-    fig, ax = plt.subplots()
-    # this locator puts ticks at regular intervals
-    loc = ticker.MultipleLocator(base=0.2)
-    ax.yaxis.set_major_locator(loc)
-    plt.plot(points)
 
 # What follows comes from the nice https://github.com/emanjavacas/pie/blob/master/pie/utils.py
 @contextmanager

### 2019-05-29 boudams/encoder.py
Started working with muficode

@@ -9,7 +9,7 @@ import json
 import copy
 from operator import itemgetter
 
-from boudams.utils import mufidecode
+from mufidecode import mufidecode
 
 DEVICE = torch.device("cuda" if torch.cuda.is_available() else "cpu")
 DEFAULT_INIT_TOKEN = "<SOS>"
@@ -260,9 +260,10 @@ class LabelEncoder:
             x = x.lower()
 
         if self.remove_diacriticals:
-            x = mufidecode(x.replace("ꝯ", "9")).strip()
-
-        x = x.split("\t")
+            x: Tuple[str, ...] = mufidecode(x, join=False)
+            x = (x[:x.index("\t")], x[x.index("\t")+1:])
+        else:
+            x = x.split("\t")
 
         if self.max_len and len(x[0]) >= len(x[1]) > self.max_len:
             raise AssertionError("Data should be smaller than maximum length")
@@ -273,7 +274,7 @@ class LabelEncoder:
             x = x.lower()
 
         if self.remove_diacriticals:
-            x = mufidecode(x.replace("ꝯ", "9")).strip()
+            x = mufidecode(x, join=False)
         return x
 
     def pad_and_tensorize(
@@ -511,7 +512,7 @@ if __name__ == "__main__":
                if l != len(reversed_data[0]) - reversed_data[0].count(" ")
            ] == [], \
         "All element should have the same size (length : %s) if we remove the spaces" % list(map(len, reversed_data))
-
+    print(reversed_data)
     assert reversed_data == [
         ['<SOS>', 'e', ' ', 'd', 'e', 'u', 's', ' ', 't', 'u', 'n', 'e', 'i', 'r', 'e', ' ', 'e', ' ', 'p', 'l', 'u',
          'i', 'e', ' ', 'm', 'e', 'r', 'v', 'e', 'i', 'l', 'l', 'u', 's', 'e', ' ', 'a', ' ', 'c', 'e', 'l', ' ', 'j',

### 2019-06-02 boudams/cli.py
Fixed a faulty if

@@ -117,8 +117,12 @@ def template(filename):
 @click.option("--epochs", type=int, default=100, help="Number of epochs to run")
 @click.option("--batch_size", type=int, default=32, help="Size of batches")
 @click.option("--device", default="cpu", help="Device to use for the network (cuda, cpu, etc.)")
-def train(config_files, epochs, batch_size, device):
+@click.option("--debug", default=False, is_flag=True)
+def train(config_files, epochs, batch_size, device, debug):
     """ Train one or more models according to [CONFIG_FILES] JSON configurations"""
+    if debug:
+        import logging
+        logging.getLogger().setLevel(logging.DEBUG)
     for config_file in config_files:
         config = json.load(config_file)
 
@@ -134,6 +138,9 @@ def train(config_files, epochs, batch_size, device):
             lower=config["label_encoder"].get("lower", True)
         )
         vocabulary.build(train_path, dev_path, test_path, debug=True)
+        if debug:
+            from pprint import pprint
+            pprint(vocabulary.mtoi)
 
         # Get the datasets
         train_dataset: DatasetIterator = vocabulary.get_dataset(

### 2019-06-02 boudams/encoder.py

@@ -254,28 +254,19 @@ class LabelEncoder:
         :param line:
         :return:
         """
-        x = line.strip()
+        inp, out = line.strip().split("\t")
+        return tuple(self.prepare(inp)), tuple(self.prepare(out))
 
-        if self.lower:
-            x = x.lower()
-
+    def prepare(self, inp: str):
         if self.remove_diacriticals:
-            x: Tuple[str, ...] = mufidecode(x, join=False)
-            x = (x[:x.index("\t")], x[x.index("\t")+1:])
+            inp = mufidecode(inp, join=False)
+            if self.lower:
+                inp = [char.lower() for char in inp]
         else:
-            x = x.split("\t")
+            if self.lower:
+                inp = inp.lower()
 
-        if self.max_len and len(x[0]) >= len(x[1]) > self.max_len:
-            raise AssertionError("Data should be smaller than maximum length")
-        return tuple(x[0]), tuple(x[1])
-
-    def prepare(self, x: str):
-        if self.lower:
-            x = x.lower()
-
-        if self.remove_diacriticals:
-            x = mufidecode(x, join=False)
-        return x
+        return inp
 
     def pad_and_tensorize(
             self,

### 2019-06-02 boudams/trainer.py

@@ -124,7 +124,7 @@ class Scorer(object):
 
 
 class LRScheduler(object):
-    def __init__(self, optimizer, mode=PlateauModes.loss, **kwargs):
+    def __init__(self, optimizer, mode=PlateauModes.accuracy, **kwargs):
         self.lr_scheduler = optim.lr_scheduler.ReduceLROnPlateau(
             optimizer, mode=mode.value, **kwargs)  # Max because accuracy :)
         self.mode = mode
@@ -229,7 +229,7 @@ class Trainer(object):
                 )
 
                 # Run a check on saving the current model
-                self._temp_save(fid, best_valid_loss, dev_score)
+                best_valid_loss = self._temp_save(fid, best_valid_loss, dev_score)
 
                 # Advance Learning Rate if needed
                 lr_scheduler.step(dev_score)
@@ -261,9 +261,11 @@ class Trainer(object):
             except EarlyStopException:
                 print("Reached plateau for too long, stopping.")
 
-        self._temp_save(fid, best_valid_loss, dev_score)
+        best_valid_loss = self._temp_save(fid, best_valid_loss, dev_score)
+
         try:
             self.tagger.model.load_state_dict(torch.load(fid))
+            print("Saving model with loss %s " % best_valid_loss)
             os.remove(fid)
         except FileNotFoundError:
             print("No model was saved during training")

### 2019-06-03 boudams/trainer.py
Early stop fixed

@@ -128,13 +128,22 @@ class LRScheduler(object):
         self.lr_scheduler = optim.lr_scheduler.ReduceLROnPlateau(
             optimizer, mode=mode.value, **kwargs)  # Max because accuracy :)
         self.mode = mode
+        self.steps = 0
 
     def step(self, score):
+        scheduler_steps = self.lr_scheduler.num_bad_epochs
         self.lr_scheduler.step(getattr(score, self.mode.name))
+        # No change in number of bad epochs =
+        #   we are progressing
+        if scheduler_steps == self.lr_scheduler.num_bad_epochs:
+            self.steps = 0
+        # Otherwise, we are not
+        else:
+            self.steps += 1
 
-    @property
-    def steps(self):
-        return self.lr_scheduler.num_bad_epochs
+        if self.steps >= self.patience * 2:
+            # If we haven't progressed even by lowering twice
+            raise EarlyStopException("No progress for %s , stoping now... " % self.steps)
 
     @property
     def patience(self):
@@ -230,10 +239,6 @@ class Trainer(object):
 
                 # Run a check on saving the current model
                 best_valid_loss = self._temp_save(fid, best_valid_loss, dev_score)
-
-                # Advance Learning Rate if needed
-                lr_scheduler.step(dev_score)
-
                 print(f'\tTrain Loss: {train_score.loss:.3f} | Perplexity: {train_score.perplexity:7.3f} | '
                       f' Acc.: {train_score.accuracy:.3f} | '
                       f' Lev.: {train_score.leven:.3f} | '
@@ -245,6 +250,9 @@ class Trainer(object):
                       f' Lev. / char: {dev_score.leven_per_char:.3f}')
                 print(lr_scheduler)
                 print()
+
+                # Advance Learning Rate if needed
+                lr_scheduler.step(dev_score)
 
                 if lr_scheduler.steps >= lr_patience and lr_scheduler.lr < min_lr:
                     raise EarlyStopException()
@@ -260,6 +268,7 @@ class Trainer(object):
                 break
             except EarlyStopException:
                 print("Reached plateau for too long, stopping.")
+                break
 
         best_valid_loss = self._temp_save(fid, best_valid_loss, dev_score)
 

### 2019-06-03 train.sh

@@ -1,0 +1,4 @@
+venv/bin/boudams train --device cuda --epochs 50 *.json
+venv/bin/boudams train --device cuda --epochs 50 *.json
+
+

### 2019-06-03 article/article.tex
Working on article, huge enhancements of some metrics

@@
 
-\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French and Latin}
+\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French}
 \author[1]{Thibault Clérice}
@@
 
+\pgfplotsset{compat=1.15}
 \begin{document}
@@
 
-\keywords{convolutional network; scripta continua; tokenization; Latin; Old French; word segmentation}
+\keywords{convolutional network; scripta continua; tokenization; Old French; word segmentation}
 
@@
 % To Read : Stutzmann article.
+% Parler plus de la variation orthographique de langue
 
@@
 \textit{Scripta continua} is a writing phenomenon where words would not be separated by spaces and it appears to have disappeared around the 8th century (see \citet{zanna1998lecture}).
-Although, spacing can be somewhat erratic in later centuries (\textit{cf.} Figure \ref{fig:4lines}).
+Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century.
 In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.
@@
 
+Indeed, while Chinese's issue seems to lie in the decomposition of relatively fix characters, Old French or medieval latin present heavy variation of spelling.
+In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spelling of the word \textit{cheval} (horse in Old and Modern French) whose apparition counts span from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}.
+This  makes a dictionary approach rather difficult as it would rely on a high number of different spelling and makes the computation highly complex.
+
 \section{Description and evaluation}
@@
 The model is based on traditional text input encoding where each character is transcoded to an index.
-Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content.
+Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content (\textit{cf.} Table \ref{lst:input_output_example}.
 
-\begin{table}[]
+\begin{table}
 \centering
@@
 
-For evaluation purposes, and to reduce the number of input classes, we propose both a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
+For evaluation purposes, and to reduce the number of input classes, we propose two options for data transcoding: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
 \ref{fig:normalization}).
-On this point, a lot of issues were found with transliteration of MUFI characters that were part of the original datasets, as they are badly interpreted by the \texttt{unidecode} python package.
+On this point, a lot of issues were found with transliteration of medieval paelographic characters that were part of the original datasets, as they are badly interpreted by the \texttt{unidecode} python package.
 Indeed, \texttt{unidecode} will simply remove characters it does not understand.
-An addendum to said package that would transcribe ligatures, abbreviations, etc.
-would probably be a good feature.
+I built a secondary package named \texttt{mufidecode} (\citet{thibault_clerice_2019_3237731}) which precedes unidecode equivalency tables when the data is known of the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).
 
-% Need a better system that this/
 \begin{figure}
   \centering
-  \includegraphics[width=\linewidth]{inputs.png}
-  \caption{ Different possibilities of pre-processing. In the last, the first character (LATIN ABBREVIATION SIGN SMALL ET WITH STROKE) is lost by \texttt{unidecode}}
+  \includegraphics[width=\linewidth]{carbon.png}
+  \caption{Different possibilities of pre-processing. The option with join=False was kept, as it keeps abbreviation marked as single characters. Note how \texttt{unidecode} loses the P WITH BAR}
   \label{fig:normalization}
@@
 
-Each model we have used is composed of an Encoder, generally built around an embedding layer, and the same Linear Classifier. The encoders tested are:
+Aside from normalizations of the input and output, three different structure of models were tested.
+Every model is composed by one encoder described below and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary).
+For final scores, SOS, EOS and PAD were ignored.
+
+The encoders are the following (configurations in parenthesis):
 
 \begin{itemize}
-  \item LSTM encoder with hidden cell
-  \item Convolutional (CNN) encoder with position embeddings 
-  \item Convolutional (CNN) encoder without position embeddings % Currently not really because it's part of the network, but it should be for the release of the article
+  \item LSTM encoder with hidden cell (Embedding (512), Dropout(0.5), Hidden Dimension (512), Layers(10))
+  \item Convolutional (CNN) encoder with position embeddings (Embedding (256), Embedding(Maximum Sentence Size=150), Kernel Size (5), Dropout(0.25), Layers (10))
+  \item Convolutional (CNN) encoder without position embeddings (Embedding (256), Kernel Size (5), Dropout(0.25), Layers (10))
 \end{itemize}
@@
 
-Datasets are transcription from manuscripts with unresolved abbreviation coming from different projects
+Datasets are transcription from manuscripts with unresolved abbreviation coming from different projects.
+The \textbf{Old French} is based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and \citet{tnah_transcription}.
+It contains
 
-\begin{itemize}
-    \item \textbf{Old French} based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and transcription from the TNAH master at the École Nationale des Chartes.
 \begin{itemize}
@@
 \end{itemize}
-\end{itemize}
+
+The input was generated by grouping at least 2 words and a maximum of 8 words together per sample.
+On a probability of 0.2, noise character could be added (noise character was set to DOT ('.')) and some words were kept randomly from a sample to another on a probability of 0.3 and a maximum number of word kept of 1.
+If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain.
+A maximum input size of 100 was kept.
+The results corpora should be varied in sizes as shown by \ref{fig:word_sizes}.
+The corpora is composed by 193 different characters when not normalized, in which some MUFI characters appears few hundred times \ref{tab:mufi_examples}.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{length.png}
+  \caption{Distribution of word size over the train, dev and test corpora}
+  \label{fig:word_sizes}
+\end{figure}
+
+\begin{table}
+\begin{tabular}{llll}
+\hline
+                                                   & Train dataset & Dev dataset & Test dataset \\ \hline
+TIRONIAN SIGN ET                                   & 4367          & 541         & 539          \\
+CON                             & 508           & 70          & 76           \\
+P WITH STROKE THROUGH DESCENDER & 580           & 69          & 84           \\ \hline
+\end{tabular}
+  \caption{Examples of some MUFI characters distributions}
+  \label{tab:mufi_examples}
+\end{table}
 
@@
 
-\begin{figure}[h!]
+The training parameters was 0.00005 in learning rate for each CNN model and 0.001 for the LSTM one, and 64 in batch sizes.
+Training reached a plateau fairly quickly for each model (\textit{cf.} \ref{fig:loss}).
+Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores})
+
+\begin{figure}
   \begin{center}
@@
         ]
-        \addplot table[x=Epoch,y=CNN Train,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN No Pos,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN lower,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN norm,col sep=comma] {accuracies.csv}; 
-        \legend{CNN Train,CNN without Position embedding, CNN with lowercase transform, CNN with ASCIIification}
+        \addplot table[x=Epoch,y=CNN 1,col sep=comma] {accuracies.csv};
+        \addplot table[x=Epoch,y=CNN L,col sep=comma] {accuracies.csv};
+        \addplot table[x=Epoch,y=CNN w/o P,col sep=comma] {accuracies.csv};
+        \addplot table[x=Epoch,y=CNN N,col sep=comma] {accuracies.csv};
+        \addplot table[x=Epoch,y=CNN L N,col sep=comma] {accuracies.csv};
+        \legend{CNN, CNN L, CNN P, CNN N, CNN L N}
       \end{axis}
     \end{tikzpicture}
-    \caption{Training Accuracy over 50 epochs}
-  \label{fig:accuracy}
+    \caption{Training Loss (Cross-entropy) until plateau was reached. N = normalized, L = Lower, P = no position embedding. LSTM was removed as it did not go below 0.65}
+  \label{fig:loss}
   \end{center}
@@
 
-% Include evaluation on test sets
+\begin{table}
+\centering
+\begin{tabular}{lllll}
+\hline
+Model   & Accuracy & Precision & Recall & FScore \\ \hline
+CNN     & 0.991    & 0.985     & 0.990  & 0.987  \\
+CNN L   & 0.991    & 0.979     & 0.990  & 0.985  \\
+CNN P   & \textbf{0.993}    & \textbf{0.990}& \textbf{0.991}  & \textbf{0.990}  \\
+CNN N   & 0.991    & 0.987     & 0.988  & 0.988  \\
+CNN L N & 0.992    & 0.988     & 0.989  & 0.988  \\
+LSTM    & 0.741    & 0.184     & 0.500  & 0.269  \\ \hline
+\end{tabular}
+\caption{Scores over the test dataset. N = normalized, L = Lower, P = no position embedding.}
+\label{tab:scores}
+\end{table}
 
@@
 
-The following input was not seen during training, dev or test, and is part of a different corpus :
+The following inputs has been tagged with the CNN P model.
+Batch are constructed around the regex \texttt{\\W} with package \texttt{regex}.
+This explains why inputs such as \texttt{".i."} are automatically tagged as \texttt{" .
+i .
+"} by the tool.
+The input was stripped of its spaces before tagging, we only show the ground truth by commodity.
 
@@
 
-\section{Annex 1}
-Pellentesque dignissim ultrices fringilla. Vivamus eu luctus ante, vel bibendum magna. Curabitur elit purus, tincidunt non dui
-vitae, elementum bibendum neque. Curabitur ullamcorper sit amet justo at hendrerit. Fusce ut arcu imperdiet nibh mollis
-tempus a aliquet tellus. Quisque pharetra cursus nisi, vel lobortis ante consectetur et. Vivamus sed congue neque. Proin
-pellentesque risus nec dui consequat rutrum. Vestibulum nunc diam, placerat quis auctor vel, faucibus non justo. Etiam
-dictum purus neque. Phasellus imperdiet mauris ligula, eu laoreet nisi elementum ut. Sed sed porta massa. Aenean faucibus
-risus ultrices ornare porta. Quisque faucibus ante a tincidunt vestibulum. Lorem ipsum dolor sit amet, consectetur adipiscing
-elit.
+\section{Annex 1 : Confusion of CNN without position embeddings}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{confusion.png}
+  \caption{Confusion matrix of the CNN model without position embedding}
+  \label{fig:confusion_matrix}
+\end{figure}
 

### 2019-06-03 article/code_excerpt.py

@@ -1,0 +1,9 @@
+import mufidecode
+import unidecode
+"sot la gnͣt abstinence dess eintes uirges ele  ꝑla"
+mufidecode.mufidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla")
+# ' sot la gnat abstinence dess eintes uirges ele  pla'
+mufidecode.mufidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla", join=False
+# (' ',  's',  'o',  't',  ' ',  'l',  'a',  ' ',  'g',  'n',  'a',  't',  ' ',  'a',  'b',  's',  't',  'i',  'n',  'e',  'n',  'c',  'e',  ' ',  'd',  'e',  's',  's',  ' ',  'e',  'i',  'n',  't',  'e',  's',  ' ',  'u',  'i',  'r',  'g',  'e',  's',  ' ',  'e',  'l',  'e',  ' ',  ' ',  'p',  'l',  'a')
+unidecode.unidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla")
+# ' sot la gnat abstinence dess eintes uirges ele  la'

### 2019-06-03 boudams/cli.py

@@ -34,7 +34,7 @@ def dataset():
 @click.option("--noise_char", type=str, default=".", help="Character to add between words for noise purposes")
 @click.option("--noise_char_random", type=float, default=0.2, help="Probability to add [NOISE_CHAR] in between words")
 @click.option("--max_noise_char", type=int, default=2, help="Maximum amount of [NOISE_CHAR] to add sequentially")
-def generate(method, output_path, input_path, min_words, max_words, min_char_length,
+def convert(method, output_path, input_path, min_words, max_words, min_char_length,
              max_char_length, random_keep, max_kept, noise_char, noise_char_random, max_noise_char):
     """ Build sequence training data using files with [METHOD] format in [INPUT_PATH] and saving the
     converted format into [OUTPUT_PATH]
@@ -58,6 +58,49 @@ def generate(method, output_path, input_path, min_words, max_words, min_char_len
         )
 
 
+@dataset.command("statistics")
+@click.argument("png_path", type=click.Path(file_okay=True, dir_okay=False))
+@click.argument("char_count", type=click.Path(file_okay=True, dir_okay=False))
+@click.argument("input_path", nargs=-1, type=click.File(mode="r"))
+def sizes(png_path, char_count, input_path):
+    """ Build sequence training data using files with [METHOD] format in [INPUT_PATH] and saving the
+    converted format into [OUTPUT_PATH]
+
+    If you are using `tsv-header` as a method, columns containing tokens should be named "tokens" or "form"
+    """
+    import matplotlib.pyplot as plt
+    from collections import Counter
+    lengths = []
+    counter = {
+        file.name: Counter()
+        for file in input_path
+    }
+    print(counter)
+    for file in input_path:
+        for l in file.readlines():
+            x, truth = l.split("\t")
+            words = truth.strip().split()
+            lengths += [len(t) for t in words]
+            counter[file.name].update(Counter(list(x)))
+
+    fig1, ax1 = plt.subplots()
+    ax1.set_title('Distribution of word sizes in the dataset')
+    ax1.boxplot(lengths)
+    plt.savefig(png_path)
+
+    words = list(set(keys for cnter in counter.values() for keys in cnter))
+    import csv
+    with open(char_count, "w") as f:
+        writer = csv.DictWriter(f, fieldnames=list(["char"] + [c for c in counter]))
+        writer.writeheader()
+        for word in sorted(words):
[…]

### 2019-06-03 boudams/model/linear.py

@@ -27,7 +27,91 @@ class LinearEncoderCNN(CNNEncoder):
 
 
 class LinearLSTMEncoder(LSTMEncoder):
-    """ Linear version of the LSTMEncoder """
+    """ Linear
+     version of the LSTMEncoder """
+
+
+class LinearEncoderCNNNoPos(nn.Module):
+    def __init__(self, input_dim, emb_dim, hid_dim, n_layers, kernel_size, dropout, device: str = "cpu"):
+        super().__init__()
+
+        assert kernel_size % 2 == 1, "Kernel size must be odd!"
+
+        self.input_dim = input_dim
+        self.emb_dim = emb_dim
+        self.hid_dim = hid_dim
+        self.kernel_size = kernel_size
+        self.dropout = dropout
+        self.device = device
+
+        self.scale = torch.sqrt(torch.FloatTensor([0.5])).to(self.device)
+
+        self.tok_embedding = nn.Embedding(input_dim, emb_dim)
+
+        self.emb2hid = nn.Linear(emb_dim, hid_dim)
+        self.hid2emb = nn.Linear(hid_dim, emb_dim)
+
+        self.convs = nn.ModuleList([nn.Conv1d(in_channels=hid_dim,
+                                              out_channels=2 * hid_dim,
+                                              kernel_size=kernel_size,
+                                              padding=(kernel_size - 1) // 2)
+                                    for _ in range(n_layers)])
+
+        self.dropout = nn.Dropout(dropout)
+
+    def forward(self, src):
+        # create position tensor
+
+        # pos = [src sent len, batch size] (Not what is documented)
+        pos = torch.arange(0, src.shape[1]).unsqueeze(0).repeat(src.shape[0], 1).to(self.device)
+
+        # embed tokens and positions
+        tok_embedded = self.tok_embedding(src)
+
+        # tok_embedded = pos_embedded = [batch size, src sent len, emb dim]
+
+        # combine embeddings by elementwise summing
+        embedded = self.dropout(tok_embedded)
+
+        # embedded = [batch size, src sent len, emb dim]
+
+        # pass embedded through linear layer to go through emb dim -> hid dim
+        conv_input = self.emb2hid(embedded)
+
+        # conv_input = [batch size, src sent len, hid dim]
+
+
+        # permute for convolutional layer
+        conv_input = conv_input.permute(0, 2, 1)
+
+        # conv_input = [batch size, hid dim, src sent len]
+
+        for i, conv in enumerate(self.convs):
+            # pass through convolutional layer
[…]

### 2019-06-03 boudams/tagger.py

@@ -85,6 +85,15 @@ class Seq2SeqTokenizer:
                     dropout=self.enc_dropout
                 )
             in_features = self.enc_hid_dim
+        elif self.system.endswith("-conv-no-pos"):
+            self.enc: linear.LinearEncoderCNNNoPos = linear.LinearEncoderCNNNoPos(
+                    self.vocabulary_dimension, emb_dim=self.emb_enc_dim,
+                    n_layers=self.enc_n_layers, hid_dim=self.enc_hid_dim,
+                    dropout=self.enc_dropout,
+                    device=self.device,
+                    kernel_size=self.enc_kernel_size
+                )
+            in_features = self.emb_enc_dim
         else:
             self.enc: linear.CNNEncoder = linear.LinearEncoderCNN(
                     self.vocabulary_dimension, emb_dim=self.emb_enc_dim,

### 2019-06-03 boudams/trainer.py

@@ -19,7 +19,7 @@ import torch.optim as optim
 import tqdm
 
 
-from sklearn.metrics import accuracy_score
+from sklearn.metrics import accuracy_score, confusion_matrix, precision_recall_fscore_support
 from leven import levenshtein
 
 from boudams.tagger import Seq2SeqTokenizer, DEVICE
@@ -47,7 +47,13 @@ class Scorer(object):
     """
     Accumulate predictions over batches and compute evaluation scores
     """
-    def __init__(self, tagger: Seq2SeqTokenizer, masked: bool = False):
+    def __init__(self, tagger: Seq2SeqTokenizer, masked: bool = False, record: bool = False):
+        """
+
+        :param tagger: Tagger
+        :param masked: Mask with Word Boundary
+        :param record: Record all inputs
+        """
         self.hypotheses = []
         self.targets = []
         self.tagger: Seq2SeqTokenizer = tagger
@@ -56,17 +62,51 @@ class Scorer(object):
         self.preds = []
         self.srcs = []
 
-        self._score_tuple = namedtuple("scores", ["accuracy", "leven", "leven_per_char"])
+        self._score_tuple = namedtuple("scores", ["accuracy", "leven", "leven_per_char",
+                                                  "precision", "recall", "fscore"])
         self.scores = None
         self.masked: bool = masked
 
+    def plot_confusion_matrix(self, path: str = "confusion-matrix.png"):
+        try:
+            from .utils import plot_confusion_matrix, plt
+        except ImportError:
+            print("You need to install matplotlib for this feature")
+            raise
+        unrolled_trues = list([y_char for y_sent in self.trues for y_char in y_sent])
+        unrolled_preds = list([y_char for y_sent in self.preds for y_char in y_sent])
+
+        plot_confusion_matrix(
+            unrolled_trues,
+            unrolled_preds,
+            labels=[self.tagger.vocabulary.space_token_index, self.tagger.vocabulary.mask_token_index],
+            classes=["WordBoundary", "WordContent"]
+            if self.tagger.vocabulary.space_token_index < self.tagger.vocabulary.mask_token_index else
+            ["WordContent", "WordBoundary"]
+        )
+        plt.savefig(path)
+
     def compute(self):
-        accuracy = [
-            accuracy_score(t, p)
-            for t, p in zip(self.trues, self.preds)
-        ]
         levenshteins = []
         leven_per_char = []
+
[…]

### 2019-06-03 boudams/utils.py

@@ -8,6 +8,11 @@ import shutil
 import unidecode
 import warnings
 
+import numpy as np
+import matplotlib.pyplot as plt
+
+from sklearn.metrics import confusion_matrix
+from sklearn.utils.multiclass import unique_labels
 
 Cache = {}
 
@@ -110,3 +115,52 @@ def ensure_ext(path, ext, infix=None):
         path = '.'.join([path, oldext])
 
     return '.'.join([path, ext])
+
+
+def plot_confusion_matrix(y_true, y_pred, classes,
+                          normalize=False,
+                          title=None,
+                          cmap=plt.cm.Blues,
+                          labels= None):
+    """
+    This function prints and plots the confusion matrix.
+    Normalization can be applied by setting `normalize=True`.
+    """
+    if not title:
+        if normalize:
+            title = 'Normalized confusion matrix'
+        else:
+            title = 'Confusion matrix, without normalization'
+
+    # Compute confusion matrix
+    cm = confusion_matrix(y_true, y_pred, labels=labels or None)
+
+    if normalize:
+        cm = cm.astype('float') / cm.sum(axis=1)[:, np.newaxis]
+
+    fig, ax = plt.subplots()
+    im = ax.imshow(cm, interpolation='nearest', cmap=cmap)
+    ax.figure.colorbar(im, ax=ax)
+    # We want to show all ticks...
+    ax.set(xticks=np.arange(cm.shape[1]),
+           yticks=np.arange(cm.shape[0]),
+           # ... and label them with the respective list entries
+           xticklabels=classes, yticklabels=classes,
+           title=title,
+           ylabel='True label',
+           xlabel='Predicted label')
+
+    # Rotate the tick labels and set their alignment.
+    plt.setp(ax.get_xticklabels(), rotation=45, ha="right",
+             rotation_mode="anchor")
+
+    # Loop over data dimensions and create text annotations.
+    fmt = '.2f' if normalize else 'd'
+    thresh = cm.max() / 2.
+    for i in range(cm.shape[0]):
+        for j in range(cm.shape[1]):
+            ax.text(j, i, format(cm[i, j], fmt),
+                    ha="center", va="center",
+                    color="white" if cm[i, j] > thresh else "black")
+    fig.tight_layout()
+    return ax

### 2019-06-03 article/article.tex
First draft of the article

@@
 \usepackage{pgfplots}
+\usepackage{tabularx}
+\newcolumntype{C}{>{\arraybackslash}X} % centered "X" column
 
@@
 % To Read : Stutzmann article.
-% Parler plus de la variation orthographique de langue
 
@@
 
-\begin{figure}
+\begin{figure}[!ht]
   \centering
@@
 
-\begin{table}
+\begin{table}[!ht]
 \centering
@@
 
-\begin{table}
+\begin{table}[!ht]
 \begin{tabular}{llll}
@@
 
-\begin{figure}
+\begin{figure}[!ht]
   \begin{center}
@@
 The following inputs has been tagged with the CNN P model.
-Batch are constructed around the regex \texttt{\\W} with package \texttt{regex}.
+Batch are constructed around the regular expression \texttt{\\W} with package \texttt{regex}.
 This explains why inputs such as \texttt{".i."} are automatically tagged as \texttt{" .
@@
 
-\begin{itemize}
-    \item \texttt{Truth :} Aies joie et leesce en ton cuer car tu auras une fille qui aura .i. fil qui sera de molt grant merite devant Dieu et de grant los entre les homes.
-    \item \texttt{Input :} Aiesjoieetleesceentoncuercartuaurasunefillequiaura.i.filquiserademoltgrantmeritedevantDieuetdegrantlosentreleshomes.
-    \item \texttt{CNN :} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
-    \item \texttt{CNN lower:}	Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
-    \item \texttt{CNN without position:} Aiesjoie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
-    \item \texttt{CNN Normalize:}	Aies joie et leesce en ton cuer car tu auras une fille qui aura . i . fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
-\end{itemize}
+\begin{table}[!ht]
+\centering
+\begin{tabularx}{\textwidth}{|C|C|}
+\hline
+\textbf{Ground truth} & \textbf{Tokenized output} \\\hline
+Aies joie et leesce en ton cuer car tu auras une fille qui aura .i.
+fil qui sera de molt grant merite devant Dieu et de grant los entre les homes.Conforte toi et soies liee car tu portes en ton ventre .i.
+fil qui son lieu aura devant Dieu et qui grant honnor fera a toz ses parenz.
+& Aies joie et leesce en ton cuer car tu auras une fille qui aura .
+i .
+fil qui sera de molt grant merite devant Dieu et de grant los entre les homes .
+Confort e toi et soies liee car tu portes en ton ventre .
+i .
+fil qui son lieu aura devant Dieu et qui grant honnor fera a toz ses parenz .
+\\\hline
+\end{tabularx}
+\caption{Output examples on a text from outside the dataset}
+\label{tab:example_output}
+\end{table}
 
@@
 
-We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view
+We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view and as a NLP task.
+Word segmentation for some text can be even difficult for humanist, and as such, we believe that post-processing of OCR through tools like Boudams can be a better way to achieve data-mining of the dataset.
+In light of the high accuracy of the model, we believe the model should perform the same way independently from the language in Medieval Western Europe.
 
-% "The purpose of the discussion is to interpret and describe the significance of your findings in light of what was already known about the research problem being investigated and to explain any new understanding or insights that emerged as a result of your study of the problem.
-The discussion will always connect to the introduction by way of the research questions or hypotheses you posed and the literature you reviewed, but the discussion does not simply repeat or rearrange the first parts of your paper; the discussion clearly explain how your study advanced the reader's understanding of the research problem from where you left them at the end of your review of prior research."
+We were surprised by the negligible effects of the different normalization methods (lower-casing; ASCII reduction; both).
+The presence of certain MUFI characters might provide enough information about segmentation and be in enough numbers for them not to impact the network weights.
 
@@
 
-While
-% The conclusion is intended to help the reader understand why your research should matter to them after they have finished reading the paper.
-A conclusion is not merely a summary of the main topics covered or a re-statement of your research problem, but a synthesis of key points and, if applicable, where you recommend new areas for future research.
-For most college-level research papers, one or two well-developed paragraphs is sufficient for a conclusion, although in some cases, three or more paragraphs may be required.
+Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more important data mining of OCRed manuscript.
+In aftermath, we wonder if the importance of normalization and lowering should be higher depending on the size of the corpora and its content.
 
-\subsection{Acknowledgement}
+\subsection{Acknowledgements}
 
@@
 
-\section{Annex 2}
-Cras tristique vel nisi at aliquet. Proin egestas erat sit amet velit lobortis imperdiet. Integer et arcu sapien. Etiam id blandit
-sapien. Nam tempus lacus ac massa semper, vel laoreet turpis rutrum. Mauris eget nibh vitae justo porta imperdiet sed vel
-ligula. In imperdiet, augue vel condimentum convallis, neque augue imperdiet neque, eget dapibus nunc mauris ultricies
-tortor. Nam eget nunc egestas, blandit lectus non, aliquam nunc. Cras sed quam vitae arcu ornare lobortis. Ut ut lacus
-hendrerit, convallis orci sit amet, commodo nunc. Pellentesque eget tincidunt tortor. Nunc ornare molestie mauris id vehicula.
-Suspendisse pharetra tortor metus, sit amet fermentum tellus vehicula ut.
-
 \end{document}

### 2019-06-04 baselines/ws.py
Added baseline with ngrams

@@ -1,0 +1,112 @@
+import os
+from collections import Counter
+from typing import Tuple
+import json
+
+import wordsegment as ws
+
+json_cache = os.path.join(os.path.dirname(__file__), "ws_cache.json")
+
+
+# Generate N-Gram pairs
+def pairs(iterable):
+    iterator = iter(iterable)
+    values = [next(iterator)]
+    for value in iterator:
+        values.append(value)
+        yield ' '.join(values)
+        del values[0]
+
+
+# The segmenter requires a clean up function. Normaly, it strips ponctuation
+# We overwrite the clean-up by not cleaning up
+def identity(value):
+    return value
+
+
+# Integer classes for Word Content (x) and Word Boundary (S)
+classes = {
+    "x": 0,
+    "S": 1
+}
+
+
+def match(segmenter: ws.Segmenter, x: str, truth: str) -> Tuple[Tuple[int, int]]:
+    """ Given a segmenter, compute the prediction and returns the integer class mask for truth and prediction
+
+    """
+    pred = segmenter.segment(x)
+
+    truth_mask = [
+        "x" if ngram[1] != " " else "S"
+        for ngram in zip(*[truth[i:] for i in range(2)])
+        if ngram[0] != " "
+    ] + ["S"]
+    pred_mask = [
+        char
+        for tok in pred
+        for char in ["x"] * (len(tok) - 1) + ["S"]
+    ]
+
+    return tuple(
+        (classes[t], classes[p])
+        for t, p in zip(truth_mask, pred_mask)
+    )
+
+# Load datasets
+text = []
+with open(os.path.join(os.path.dirname(__file__), "..", "datasets", "fro", "train.tsv")) as input_text:
+    for line in input_text.readlines():
+        inp, truth = tuple(line.strip().lower().split("\t"))
+        text.extend(truth.split())
+
+# Change the segmenter information
+segmenter = ws.Segmenter()
+
+# Generate the dictionary and set-up the segmenter
+segmenter.unigrams.clear()
+segmenter.unigrams.update(Counter(text))
+segmenter.bigrams.clear()
+segmenter.bigrams.update(Counter(pairs(text)))
+segmenter.words = list(set(text))
+
+
+segmenter.total = float(sum(segmenter.unigrams.values()))
+segmenter.clean = identity
+segmenter.limit = max(list(map(len, text)))
+
+# Load the test corpus
+test = []
+padding = 0
+with open(os.path.join(os.path.dirname(__file__), "..", "datasets", "fro", "test.tsv")) as input_text:
[…]

### 2019-06-04 baselines/ws.py
Added an unknown corpus for baseline evaluation purposes

@@ -5,7 +5,8 @@ import json
 
 import wordsegment as ws
 
-json_cache = os.path.join(os.path.dirname(__file__), "ws_cache.json")
+corpus = "fro"
+json_cache = os.path.join(os.path.dirname(__file__), corpus+"_" + "ws_cache.json")
 
 
 # Generate N-Gram pairs
@@ -55,7 +56,7 @@ def match(segmenter: ws.Segmenter, x: str, truth: str) -> Tuple[Tuple[int, int]]
 
 # Load datasets
 text = []
-with open(os.path.join(os.path.dirname(__file__), "..", "datasets", "fro", "train.tsv")) as input_text:
+with open(os.path.join(os.path.dirname(__file__), "..", "datasets", corpus, "train.tsv")) as input_text:
     for line in input_text.readlines():
         inp, truth = tuple(line.strip().lower().split("\t"))
         text.extend(truth.split())
@@ -78,7 +79,7 @@ segmenter.limit = max(list(map(len, text)))
 # Load the test corpus
 test = []
 padding = 0
-with open(os.path.join(os.path.dirname(__file__), "..", "datasets", "fro", "test.tsv")) as input_text:
+with open(os.path.join(os.path.dirname(__file__), "..", "datasets", corpus, "test.tsv")) as input_text:
     for line in input_text.readlines():
         inp, truth = tuple(line.strip().lower().split("\t"))
         test.extend([(inp, truth)])
@@ -104,9 +105,50 @@ else:
         results = json.load(f)
 
 from sklearn.metrics import accuracy_score, precision_recall_fscore_support, confusion_matrix
-print("Accuracy:", accuracy_score(results["truth"], results["pred"]))
-print("Precision, Recall, FScore, Support", precision_recall_fscore_support(results["truth"], results["pred"], average="macro"))
-print("Matrix", confusion_matrix(results["truth"], results["pred"]))
+print("- Accuracy: ", accuracy_score(results["truth"], results["pred"]))
+p, r, f, _ = precision_recall_fscore_support(results["truth"], results["pred"], average="macro")
+print("- Precision: ", p)
+print("- Recall: ", r)
+print("- FScore: ", f)
+print("- Matrix", confusion_matrix(results["truth"], results["pred"]))
 
 with open(os.path.join(os.path.dirname(__file__), "..", "article", "input.txt")) as input_text:
-    print("Completely unknown text", " ".join(segmenter.segment(input_text.read())))
[…]

### 2019-06-04 generate.sh

@@ -1,9 +1,12 @@
 #!/usr/bin/env bash
 
-venv/bin/boudams dataset convert plain-text test_data/bfmmss/gt test_data/bfmmss/txt/*.txt
-venv/bin/boudams dataset convert plain-text test_data/geste/gt test_data/geste/src/*.txt
-venv/bin/boudams dataset convert plain-text test_data/bnf412/gt test_data/bnf412/txt/*.txt
-venv/bin/boudams dataset convert plain-text test_data/himanis/gt/fro test_data/himanis/fro/*.txt
-venv/bin/boudams dataset convert plain-text test_data/pinche/gt test_data/pinche/src/*.txt
+# Params :
+# --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/bfmmss/gt test_data/bfmmss/txt/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/geste/gt test_data/geste/src/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/bnf412/gt test_data/bnf412/txt/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/himanis/gt/fro test_data/himanis/fro/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/pinche/gt test_data/pinche/src/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
 
-venv/bin/boudams dataset generate datasets/fro test_data/bfmmss/gt/*.txt test_data/geste/gt/*.txt test_data/bnf412/gt/*.txt test_data/himanis/gt/fro/*.txt test_data/pinche/gt/*.txt
+#  --max_char_length 250
[…]

### 2019-06-04 generate_unknown.sh

@@ -1,0 +1,4 @@
+#!/usr/bin/env bash
+
+venv/bin/boudams dataset convert plain-text test_data/unknown/gt test_data/unknown/*.txt --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 140 --random_keep 0 --noise_char_random 0
+venv/bin/python test_data/unknown/consolidate.py

### 2019-06-04 test_data/unknown/consolidate.py

@@ -1,0 +1,10 @@
+import glob
+import os
+
+here = os.path.join(os.path.dirname(__file__))
+
+with open(os.path.join(here, "..", "..", "datasets", "fro", "unknown.tsv"), "w") as output:
+    for file in glob.glob(os.path.join(here, "gt", "*.txt")):
+        with open(file) as input:
+            for line in input.readlines():
+                output.write(line)

### 2019-06-04 article/article.tex
Updated of the article

@@
 \abstract{Tokenization of modern and old Western European languages seems to be fairly simple as it stands on the presence mostly of markers such as spaces and punctuation.
-Although, when dealing with old sources like manuscript written in \textit{scripta continua}, (1) such markers are mostly absent, (2) spelling variation and rich morphology makes dictionary based approaches difficult.
+Although, when dealing with old sources like manuscript written in \textit{scripta continua} or later manuscripts, (1) such markers are mostly absent, (2) spelling variation and rich morphology makes dictionary based approaches difficult.
 We show that applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence can be used to tokenize such inputs.
@@
 Training reached a plateau fairly quickly for each model (\textit{cf.} \ref{fig:loss}).
-Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores})
+Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores}).
+To compare the results, we used the \texttt{wordsegment} package \citet{WordSegment} as a baseline.
 
@@
 Model    & Accuracy & Precision & Recall & FScore \\ \hline
+Baseline & 0.989    & 0.986     & 0.984  & 0.985 \\
 CNN      & 0.991    & 0.985     & 0.990  & 0.987  \\
@@
 \label{tab:scores}
+\end{table}
+
+\subsubsection{Unknown texts}
+
+While all model using CNN shows improvement over the baseline, the model definitely does not outperform it by a huge margin (\textless 0.02 FScore).
+And for a reason : the baseline already performs nearly perfectly on the test corpus.
+The dictionary attack using n-grams did actually perform well.
+As a result, we wanted to compare how both models would perform on a secondary test corpus composed by texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same inputs.
+As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the Graal \citet{graal}, a Passion and a Vie de Saint Leger \citet{old_french_corpus}, a Vie de Saint Thibaut \citet{theobaldus}.
+No noise characters and no random keeping of words were applied.
+
+The results here were highly different (\textit{cf.} Table \ref{tab:scores_unknown}): while it appears that the CNN is able to expand its "comprehension" of the language to newer texts, the new words are more difficult to take into account for the baseline \texttt{wordsegment} n-gram approach, resulting in a respective drop to 0.945 and 0.838 FScore.
+WordSegment specifically performed badly with WB false positives : it had 3658 over a corpus containing 10,193 WB token (around 35 \%).
+
+\begin{table}[!ht]
+\centering
+\begin{tabular}{lllllll}
+\hline
+ & Accuracy & Precision & Recall & FScore & WB as WC & WC as WB \\ \hline
+Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & 644 \\
+CNN P & \textbf{0.957} & \textbf{0.948} & \textbf{0.944} & \textbf{0.945} & \textbf{854} & \textbf{723} \\ \hline
+\end{tabular}
+\caption{Scores over the unknown dataset. WB as WC and WC as WB are taken from the confusion matrices}
+\label{tab:scores_unknown}
 \end{table}
@@
 
+While the baseline surprised us by performing this well on the test corpus, it definitely performed less well than the CNN on a completely unknown corpus: in this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical ngram approaches.
+
 \subsection{Conclusion}

### 2019-06-04 generate_latin.sh
Added data for latin and medieval latin

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+# Params :
+# --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/latin_prose/gt test_data/latin_prose/src/**/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/latin_poetry/gt test_data/latin_poetry/src/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+
+#  --max_char_length 250
+venv/bin/boudams dataset generate datasets/latin test_data/latin_prose/gt/*.txt  --max_char_length 150
+venv/bin/python test_data/latin_poetry/consolidate.py

### 2019-06-04 generate_medieval_latin.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+# Params :
+# --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/medieval_latin/monumenta/gt test_data/medieval_latin/monumenta/src/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/medieval_latin/cartae/gt test_data/medieval_latin/cartae/src/*.normalized.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+
+#  --max_char_length 250
+venv/bin/boudams dataset generate datasets/medieval_latin test_data/medieval_latin/cartae/gt/*.txt  --max_char_length 150
+venv/bin/python test_data/medieval_latin/consolidate.py

### 2019-06-04 test_data/latin_poetry/consolidate.py

@@ -1,0 +1,10 @@
+import glob
+import os
+
+here = os.path.join(os.path.dirname(__file__))
+
+with open(os.path.join(here, "..", "..", "datasets", "latin", "unknown.tsv"), "w") as output:
+    for file in glob.glob(os.path.join(here, "gt", "*.txt")):
+        with open(file) as input:
+            for line in input.readlines():
+                output.write(line)

### 2019-06-04 test_data/medieval_latin/cartae/clean.py

@@ -1,0 +1,15 @@
+# Clean up files from the search fold of the source
+
+import glob
+import os
+
+here = os.path.join(os.path.dirname(__file__))
+sep = "******"
+
+
+for file in glob.glob(os.path.join(here, "src", "*.txt")):
+    with open(file) as f:
+        text = f.read()
+    text = text.split(sep)[-2]
+    with open(file+".normalized.txt", "w") as f:
+        f.write(text)

### 2019-06-04 test_data/medieval_latin/consolidate.py

@@ -1,0 +1,10 @@
+import glob
+import os
+
+here = os.path.join(os.path.dirname(__file__))
+
+with open(os.path.join(here, "..", "..", "datasets", "medieval_latin", "unknown.tsv"), "w") as output:
+    for file in glob.glob(os.path.join(here, "monumenta", "gt", "*.txt")):
+        with open(file) as input:
+            for line in input.readlines():
+                output.write(line)

### 2019-06-04 article/article.tex
Update of the article

@@
 \usepackage{pgfplots}
+\usepackage{float}
 \usepackage{tabularx}
@@
 Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando".
-\textit{Scripta continua} is a writing phenomenon where words would not be separated by spaces and it appears to have disappeared around the 8th century (see \citet{zanna1998lecture}).
+\textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}).
 Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century.
@@
     \item 25,512 test examples
+    \item Number of classes in testing examples: 482,776 WC; 169,094 WB
+    \item Number of classes in unknown examples: 26,393 WC; 10,193 WB
 \end{itemize}
@@
 Word segmentation for some text can be even difficult for humanist, and as such, we believe that post-processing of OCR through tools like Boudams can be a better way to achieve data-mining of the dataset.
-In light of the high accuracy of the model, we believe the model should perform the same way independently from the language in Medieval Western Europe.
 
@@
 While the baseline surprised us by performing this well on the test corpus, it definitely performed less well than the CNN on a completely unknown corpus: in this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical ngram approaches.
+In light of the high accuracy of the CNN model, we believe the model should perform the same way independently from the language in Medieval Western Europe, and results in annexes on Latin confirms this for at least 2 other corpora.
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[H]
   \centering
@@
 
+\section{Annex 2 : Scores on Latin Prose and Poetic Corpora}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{llllllll}
+\hline
+ & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
+Baseline & Test & 0.978 & 0.961 & 0.974 & 0.968 & 886 & 1893 \\
+CNN P & Test & \textbf{0.992} & \textbf{0.987} & \textbf{0.989} & \textbf{0.988} & \textbf{439} & \textbf{584} \\ \hline
+Baseline & Unknown & 0.933 & 0.897 & 0.890 & 0.893 & 1587 & 1409 \\
+CNN P & Unknown & 0.970 & 0.952 & 0.956 & 0.954 & 600 & 709 \\ \hline
+\end{tabular}
+\caption{Scores over the Latin classical datasets. FN = False Negative, FP = False Positive}
+\label{tab:latin_corpora}
+\end{table}
+
+The Latin data is much more noisy than the Old French, as it was less curated than the digital edition we found for Old French.
+They are part of the Perseus corpus \citet{perseus} and were cut into passages in the context of my thesis.
+The training, evaluation and test corpora are built upon prose work from Cicero and Suetonius.
+The unknown corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc.
+Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+
+\textbf{Statistics:}
+\begin{itemize}
+\item Number of training examples: 30725
+\item Number of evaluation examples: 3558
+\item Number of testing examples: 4406
+\item Number of classes in testing examples: 105,915 WC; 26,404 WB
+\item Number of classes in unknown examples: 35,910 WC; 8,828 WB
+\end{itemize}
+
+\section{Annex 3: Scores on Medieval Latin corpora}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{llllllll}
+\hline
+ & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
+Baseline & Test & 0.989 & 0.981 & 0.986 & 0.982 & 1036 & 933 \\
+CNN P & Test & \textbf{0.997} & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & \textbf{251} & \textbf{298} \\ \hline
+Baseline & Unknown & 0.929 & 0.900 & 0.865 & 0.881 & 14,382 & 27,019 \\
+CNN P & Unknown & 0.976 & 0.960 & 0.963 & 0.962 & 6509 & 7444\\ \hline
+\end{tabular}
+\caption{Scores over the Latin medieval datasets. FN = False Negative, FP = False Positive}
+\label{tab:medieval_latin_corpora}
+\end{table}
+
+% test : 136,429+1036 WC / 933+33,120 WB
+% unk  : 458,273+14,382 WC / 27,019+85,985 WB
+
+The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data \citet{formulae} for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
+Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.
+
+\textbf{Statistics:}
+
+\begin{itemize}
+\item Number of training examples: 36814
+\item Number of evaluation examples: 4098
+\item Number of testing examples: 5612
+\item Number of classes in testing examples: 137,465 WC; 34,053 WB
+\item Number of classes in unknown examples: 472,655 WC; 113,004 WB
+\end{itemize}
+
 \end{document}

### 2019-06-05 baselines/ws.py
Latin Epigrapy

@@ -5,7 +5,7 @@ import json
 
 import wordsegment as ws
 
-corpus = "fro"
+corpus = "latin_epigraphy"
 json_cache = os.path.join(os.path.dirname(__file__), corpus+"_" + "ws_cache.json")
 
 
@@ -86,7 +86,7 @@ with open(os.path.join(os.path.dirname(__file__), "..", "datasets", corpus, "tes
         padding = max(len(truth), padding)
 
 # If compute is True, recompute the data without using the cache
-compute = False
+compute = True
 if compute:
     results = list([
         res

### 2019-06-05 generate_latin_epigraphy.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+# Params :
+# --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
+venv/bin/boudams dataset convert plain-text test_data/edh_epidoc/gt test_data/edh_epidoc/txt/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+venv/bin/python test_data/edh_epidoc/consolidate.py
+
+#  --max_char_length 250
+venv/bin/boudams dataset generate datasets/latin_epigraphy test_data/edh_epidoc/gt.tsv  --max_char_length 150
+# venv/bin/python test_data/medieval_latin/consolidate.py

### 2019-06-05 test_data/edh_epidoc/consolidate.py

@@ -1,0 +1,10 @@
+import glob
+import os
+
+here = os.path.join(os.path.dirname(__file__))
+
+with open(os.path.join(here, "gt.tsv"), "w") as output:
+    for file in glob.glob(os.path.join(here, "gt", "*.txt")):
+        with open(file) as input:
+            for line in input.readlines():
+                output.write(line)

### 2019-06-05 test_data/edh_epidoc/convert.py

@@ -1,0 +1,33 @@
+from lxml import etree as ET
+import os
+import glob
+
+ET.FunctionNamespace("http://exslt.org/regular-expressions").prefix = 're'
+
+target = os.path.join(
+    os.path.dirname(__file__)
+)
+
+xsl_xml = ET.fromstring("""<xsl:stylesheet xmlns:xsl="http://www.w3.org/1999/XSL/Transform"
+    xmlns:xs="http://www.w3.org/2001/XMLSchema"
+    xmlns:tei="http://www.tei-c.org/ns/1.0"
+    exclude-result-prefixes="xs"
+    version="1.0">
+    <xsl:output method="text"/>
+    <xsl:template match="/">
+        <xsl:apply-templates select="//tei:div[@type='edition']" />
+    </xsl:template>
+    <xsl:template match="tei:ex" />
+    <xsl:template match="tei:head"/>
+</xsl:stylesheet>""")
+
+xsl = ET.XSLT(xsl_xml)
+
+
+for file in glob.glob(os.path.join(target, "xml", "*.xml")):
+    output = os.path.join(target, "txt", os.path.basename(file)+".txt")
+    with open(file) as io_input:
+        xml = ET.parse(io_input)
+
+        with open(output, "w") as io_output:
+            io_output.write(xsl.tostring(xsl(xml)))

### 2019-06-05 article/article.tex
Updated article

@@
 \textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}).
-Never the less, spacing can be somewhat erratic in later centuries writings, as show by Figure \ref{fig:4lines}, a document from the 13th century.
-In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution \textbf{CITE JBCAMPS ?}.
+Never the less, spacing can be somewhat erratic in later centuries writings as \citet{stutzmann} has shown (\textit{cf.} Figure \ref{fig:4lines}, a document from the 13th century, displays this kind of situation).
+In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution or simply allow full-text search.
 
@@
 As a result, we wanted to compare how both models would perform on a secondary test corpus composed by texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same inputs.
-As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the Graal \citet{graal}, a Passion and a Vie de Saint Leger \citet{old_french_corpus}, a Vie de Saint Thibaut \citet{theobaldus}.
+As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
 No noise characters and no random keeping of words were applied.
@@
 \hline
- & Accuracy & Precision & Recall & FScore & WB as WC & WC as WB \\ \hline
+ & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
 Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & 644 \\
@@
 \end{tabular}
-\caption{Scores over the unknown dataset. WB as WC and WC as WB are taken from the confusion matrices}
+\caption{Scores over the unknown dataset. FN = False Negative, FP = False Positive}
 \label{tab:scores_unknown}
@@
 
+\textbf{Example:}
+
+\begin{itemize}
+    \item Input : operecuperemdeberemqueprofecto
+    \item Output : opere cuperem deberemque profecto
+
+\end{itemize}
+
 \section{Annex 3: Scores on Medieval Latin corpora}
@@
 
-The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data \citet{formulae} for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
+The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
 Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
@@
 
+\textbf{Example:}
+
+\begin{itemize}
+    \item Input : nonparvamremtibi
+    \item Output : non parvam rem tibi
+
+\end{itemize}
+
+\section{Annex 4: Scores on Latin epigraphic corpora}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{llllllll}
+\hline
+ & Corpus & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
+Baseline & Test & 0.956 & 0.935 & 0.943 & 0.939 & 2646 & 3547 \\
+CNN P & Test & \textbf{0.987} & \textbf{0.983} & \textbf{0.979} & \textbf{0.981} & \textbf{1149} & \textbf{722} \\ \hline
+Baseline & Test Uppercase & 0.956 & 0.935 & 0.942 & 0.938 & 2664 & 3457 \\
+CNN P & Test Uppercase & \textbf{0.979} & \textbf{0.972} & \textbf{0.967} & \textbf{0.969} & \textbf{1715} & \textbf{1275} \\ \hline
+Baseline & Unknown & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
+CNN P & Unknown & 0.953 & 0.939 & 0.926 & 0.932 & 4689 & 3112 \\ \hline
+Baseline & Unknown Uppercase & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
+CNN P & Unknown Uppercase & 0.936 & 0.914 & 0.902 & 0.908 & 6152 & 4464 \\ \hline
+\end{tabular}
+\caption{Scores over the Latin epigraphic datasets. FN = False Negative, FP = False Positive}
+\label{tab:epigraphyc_latin}
+\end{table}
+
+% test : 136,429+1036 WC / 933+33,120 WB
+% unk  : 458,273+14,382 WC / 27,019+85,985 WB
+
+The epigraphic Latin corpora is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the corpus of unknown is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
+We also decided to evaluate both the baseline and the model on uppercased data, as it would normally be the state the text would be found in.
+Each of the corpora presents a high number of unresolved abbreviations (\textit{ie.} one letter words).
+Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+The data presents some polytonic Greek characters, some sample being only in Greek.
+
+While CNN is influenced by
+\textbf{Statistics:}
+
+\begin{itemize}
+\item Number of training examples: 46,423
+\item Number of evaluation examples: 5,802
+\item Number of testing examples: 5,804
+\item Number of classes in testing examples: 107,963 WC; 31,900 WB
+\item Number of classes in unknown examples: 127,268 WC; 38,055 WB
+\end{itemize}
+
+\textbf{Example:}
+
+\begin{itemize}
+    \item Input : DnFlClIuliani
+    \item Output : D n Fl Cl Iuliani
+\end{itemize}
+
 \end{document}

### 2019-06-05 baselines/ws.py
Added epigraphic data

@@ -5,7 +5,7 @@ import json
 
 import wordsegment as ws
 
-corpus = "latin_epigraphy"
+corpus = "latin_epigraphy_uppercase"
 json_cache = os.path.join(os.path.dirname(__file__), corpus+"_" + "ws_cache.json")
 
 
@@ -19,7 +19,7 @@ def pairs(iterable):
         del values[0]
 
 
-# The segmenter requires a clean up function. Normaly, it strips ponctuation
+# The segmenter requires a clean up function. Normally, it strips ponctuation
 # We overwrite the clean-up by not cleaning up
 def identity(value):
     return value
@@ -62,7 +62,12 @@ with open(os.path.join(os.path.dirname(__file__), "..", "datasets", corpus, "tra
         text.extend(truth.split())
 
 # Change the segmenter information
-segmenter = ws.Segmenter()
+class Segmenter(ws.Segmenter):
+    def clean(cls, text):
+        return identity(text)
+
+
+segmenter = Segmenter()
 
 # Generate the dictionary and set-up the segmenter
 segmenter.unigrams.clear()
@@ -73,7 +78,6 @@ segmenter.words = list(set(text))
 
 
 segmenter.total = float(sum(segmenter.unigrams.values()))
-segmenter.clean = identity
 segmenter.limit = max(list(map(len, text)))
 
 # Load the test corpus

### 2019-06-05 generate_latin_epigraphy.sh

@@ -2,5 +2,9 @@
 
 # Params :
 # --min_words 2 --max_words 25 --min_char_length 8 --max_char_length 250 --random_keep 0 --noise_char_random 0
-venv/bin/boudams dataset convert plain-text test_data/edh_epidoc/pompei_gt test_data/edh_epidoc/pompei/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
-venv/bin/python test_data/edh_epidoc/consolidate_pompei.py
+venv/bin/boudams dataset convert plain-text test_data/edh_epidoc/gt test_data/edh_epidoc/txt/*.txt --max_char_length 150 --random_keep 0 --noise_char_random 0
+venv/bin/python test_data/edh_epidoc/consolidate.py
+
+#  --max_char_length 250
+venv/bin/boudams dataset generate datasets/latin_epigraphy test_data/edh_epidoc/gt.tsv  --max_char_length 150
+# venv/bin/python test_data/medieval_latin/consolidate.py

### 2019-06-05 test_data/edh_epidoc/consolidate_pompei.py

@@ -3,8 +3,10 @@ import os
 
 here = os.path.join(os.path.dirname(__file__))
 
-with open(os.path.join(here, "gt.tsv"), "w") as output:
-    for file in glob.glob(os.path.join(here, "gt", "*.txt")):
-        with open(file) as input:
-            for line in input.readlines():
-                output.write(line)
+with open(os.path.join(here, "..", "..", "datasets", "latin_epigraphy", "unknown.tsv"), "w") as output:
+    with open(os.path.join(here, "..", "..", "datasets", "latin_epigraphy_uppercase", "unknown.tsv"), "w") as output_bis:
+        for file in glob.glob(os.path.join(here, "pompei_gt", "*.txt")):
+            with open(file) as input:
+                for line in input.readlines():
+                    output.write(line)
+                    output_bis.write(line.upper())

### 2019-06-05 test_data/edh_epidoc/convert_pompei.py

@@ -1,0 +1,33 @@
+from lxml import etree as ET
+import os
+import glob
+
+ET.FunctionNamespace("http://exslt.org/regular-expressions").prefix = 're'
+
+target = os.path.join(
+    os.path.dirname(__file__)
+)
+
+xsl_xml = ET.fromstring("""<xsl:stylesheet xmlns:xsl="http://www.w3.org/1999/XSL/Transform"
+    xmlns:xs="http://www.w3.org/2001/XMLSchema"
+    xmlns:tei="http://www.tei-c.org/ns/1.0"
+    exclude-result-prefixes="xs"
+    version="1.0">
+    <xsl:output method="text"/>
+    <xsl:template match="/">
+        <xsl:apply-templates select="//tei:div[@type='edition']" />
+    </xsl:template>
+    <xsl:template match="tei:ex" />
+    <xsl:template match="tei:head"/>
+</xsl:stylesheet>""")
+
+xsl = ET.XSLT(xsl_xml)
+
+
+for file in glob.glob(os.path.join("/home/thibault/dev/canonicals/pompei-inscriptions/data", "**", "*-lat1.xml"), recursive=True):
+    output = os.path.join(target, "pompei", os.path.basename(file)+".txt")
+    with open(file) as io_input:
+        xml = ET.parse(io_input)
+
+        with open(output, "w") as io_output:
+            io_output.write(xsl.tostring(xsl(xml)).strip())

### 2019-06-07 boudams/tagger.py
Removed position when not using them

@@ -94,6 +94,8 @@ class Seq2SeqTokenizer:
                     kernel_size=self.enc_kernel_size
                 )
             in_features = self.emb_enc_dim
+            # This model does not need sentence length
+            self.out_max_sentence_length = None
         else:
             self.enc: linear.CNNEncoder = linear.LinearEncoderCNN(
                     self.vocabulary_dimension, emb_dim=self.emb_enc_dim,
@@ -193,20 +195,17 @@ class Seq2SeqTokenizer:
         strings = ["".join(tempList[n:n + 2]) for n in range(0, len(splits), 2)]
         strings = list(filter(len, strings))
 
-        treated = []
-        max_size = self.out_max_sentence_length - 5
-        for string in strings:
-            if len(string) > max_size:
-                treated.extend([
-                    "".join(string[n:n + max_size])
-                    for n in range(0, len(string), max_size)
-                ])
-            else:
-                treated.append(string)
+        if self.out_max_sentence_length:
+            treated = []
+            max_size = self.out_max_sentence_length - 5
+            for string in strings:
+                if len(string) > max_size:
+                    treated.extend([
+                        "".join(string[n:n + max_size])
+                        for n in range(0, len(string), max_size)
+                    ])
+                else:
+                    treated.append(string)
+            strings = treated
 
-        for string in treated:
-            if len(string) > max_size:
-                print(len(string))
-                print(string)
-
-        yield from self.annotate(treated, batch_size=batch_size)
+        yield from self.annotate(strings, batch_size=batch_size)

### 2019-06-07 article/article.tex
Updated the article with style correction, moved annex into Evaluation,, etc.

@@
 
-\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French}
+\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French and Latin}
 \author[1]{Thibault Clérice}
@@
 
-\abstract{Tokenization of modern and old Western European languages seems to be fairly simple as it stands on the presence mostly of markers such as spaces and punctuation.
-Although, when dealing with old sources like manuscript written in \textit{scripta continua} or later manuscripts, (1) such markers are mostly absent, (2) spelling variation and rich morphology makes dictionary based approaches difficult.
-We show that applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence can be used to tokenize such inputs.
-Additionally, we release a software with a rather simple interface for tokenizing one's corpus.}
+\abstract{Tokenization of modern and old Western European languages seems to be fairly simple, as it stands on the presence mostly of markers such as spaces and punctuation.
+However, when dealing with old sources like manuscripts written in \textit{scripta continua}, antiquity epigraphy or Middle Age manuscripts, (1) such markers are mostly absent, (2) spelling variation and rich morphology make dictionary based approaches difficult.
+Applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence is shown to be effective at tokenizing such inputs.
+Additionally, the software is released with a simple interface for tokenizing a corpus or generating a training set.}
 
@@
 
-% To Read : Stutzmann article.
-
-Tokenization of space-less strings is a task that is specifically difficult for computer when compared to "whathumancando".
-\textit{Scripta continua} is a writing phenomenon where words are be separated by spaces and has disappeared around the 8th century (see \citet{zanna1998lecture}).
-Never the less, spacing can be somewhat erratic in later centuries writings as \citet{stutzmann} has shown (\textit{cf.} Figure \ref{fig:4lines}, a document from the 13th century, displays this kind of situation).
+Tokenization of space-less strings is a task that is specifically difficult for computers as compared to "whathumanscando".
+\textit{Scripta continua} is a writing phenomenon in which words are separated by spaces that disappeared around the 8th century (see \citet{zanna1998lecture}).
+Nevertheless, spacing can be somewhat erratic in later centuries writings as \citet{stutzmann} explains (\textit{cf.} Figure \ref{fig:4lines}) and it becomes an issue for OCR, as continuous bag of word is only interesting when those are not glued together.
 In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution or simply allow full-text search.
@@
 
-We must stress in this study that the difficulty that we face is different for \textit{scripta continua} than the ones researchers face languages such as Chinese for which an already impressive amount of work has been done as it.
-Indeed, Chinese word segmentation has lately been driven by deep learning methods, specifically ones based on \textit{sequence to sequence translations}: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses BiDirectional GRU and CRF.
-Actually, meanwhile redacting this article and producing the code-base, \citet{huang2019realistic} took the same approach of encoding to linear classification to both word boundary (WB) and word content (WC) for Chinese word segmentation.
+It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese for which an already impressive amount of work has been.
+Chinese word segmentation has lately been driven by deep learning methods, specifically ones based on \textit{sequence to sequence translations}: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses BiDirectional GRU and CRF.
+\footnote{\citet{huang2019realistic} actually gave us the denomination used here: word boundary (WB) and word content (WC).}
 
-Indeed, while Chinese's issue seems to lie in the decomposition of relatively fix characters, Old French or medieval latin present heavy variation of spelling.
-In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spelling of the word \textit{cheval} (horse in Old and Modern French) whose apparition counts span from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}.
-This  makes a dictionary approach rather difficult as it would rely on a high number of different spelling and makes the computation highly complex.
+Indeed, while the issue with Chinese seems to lie in the decomposition of relatively fixed characters, Old French or Medieval Latin present heavy variation of spelling.
+In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spellings of the word "\textit{cheval}" (horse in Old and Modern French) whose apparition counts span from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}.
+This  makes a dictionary-based approach rather difficult as it would rely on a high number of different spellings, making the computation highly complex.
 
@@
 
-For evaluation purposes, and to reduce the number of input classes, we propose two options for data transcoding: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
+For evaluation purposes, and to reduce the number of input classes, two options for data transcoding were used: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
 \ref{fig:normalization}).
-On this point, a lot of issues were found with transliteration of medieval paelographic characters that were part of the original datasets, as they are badly interpreted by the \texttt{unidecode} python package.
+On this point, a lot of issues were encountered with transliteration of medieval paelographic characters that were part of the original datasets, as they are poorly interpreted by the \texttt{unidecode} python package.
 Indeed, \texttt{unidecode} will simply remove characters it does not understand.
-I built a secondary package named \texttt{mufidecode} (\citet{thibault_clerice_2019_3237731}) which precedes unidecode equivalency tables when the data is known of the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).
+A derivative package named \texttt{mufidecode} was built for this reason(\citet{thibault_clerice_2019_3237731}): it takes precedent over \texttt{unidecode} equivalency tables when the data is known of the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).
 
@@
 
-Aside from normalizations of the input and output, three different structure of models were tested.
-Every model is composed by one encoder described below and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary).
+Aside from normalizations of the input and output, three different model structures were tested.
+Every model is composed of one encoder, as described below, and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary).
 For final scores, SOS, EOS and PAD were ignored.
@@
 
-Datasets are transcription from manuscripts with unresolved abbreviation coming from different projects.
+The dataset is composed of transcriptions (from different projects) of manuscripts with unresolved abbreviation.
 The \textbf{Old French} is based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and \citet{tnah_transcription}.
@@
 
-The input was generated by grouping at least 2 words and a maximum of 8 words together per sample.
-On a probability of 0.2, noise character could be added (noise character was set to DOT ('.')) and some words were kept randomly from a sample to another on a probability of 0.3 and a maximum number of word kept of 1.
-If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain.
-A maximum input size of 100 was kept.
-The results corpora should be varied in sizes as shown by \ref{fig:word_sizes}.
-The corpora is composed by 193 different characters when not normalized, in which some MUFI characters appears few hundred times \ref{tab:mufi_examples}.
+Examples were generated automatically.
+They are between 2 and 8 words-length.
+In order to recreate the condition of OCR noise, dots were added randomly (0.2) between words.
+In order to augment the dataset, words are randomly (0.1) passed over the next example\footnote{This data augmentation was limited to one word per sample}.
+If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain, independently of the maximum number of words.
+The example however should not go beyond 100 characters.
+The results corpora should be varied in sizes as shown by Figure \ref{fig:word_sizes}.
+The corpora contains 193 different characters when not normalized, in which some MUFI characters appears few hundred times (\textit{cf.} Table \ref{tab:mufi_examples}).
 
@@
 
-The training parameters was 0.00005 in learning rate for each CNN model and 0.001 for the LSTM one, and 64 in batch sizes.
+The training parameters were 0.00005 in learning rate for each CNN model, 0.001 for the LSTM model, and batch sizes of 64.
 Training reached a plateau fairly quickly for each model (\textit{cf.} \ref{fig:loss}).
 Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores}).
-To compare the results, we used the \texttt{wordsegment} package \citet{WordSegment} as a baseline.
+To compare the results, the \texttt{wordsegment} package \citet{WordSegment} was used as a baseline.
 
@@
 \centering
-\begin{tabular}{lllll}
+\begin{tabular}{lllllll}
 \hline
-Model    & Accuracy & Precision & Recall & FScore \\ \hline
-Baseline & 0.989    & 0.986     & 0.984  & 0.985 \\
-CNN      & 0.991    & 0.985     & 0.990  & 0.987  \\
-CNN L    & 0.991    & 0.979     & 0.990  & 0.985  \\
-CNN P    & \textbf{0.993}    & \textbf{0.990}& \textbf{0.991}  & \textbf{0.990}  \\
-CNN N    & 0.991    & 0.987     & 0.988  & 0.988  \\
-CNN L N  & 0.992    & 0.988     & 0.989  & 0.988  \\
-LSTM     & 0.741    & 0.184     & 0.500  & 0.269  \\ \hline
+Model    & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
+Baseline & 0.989    & 0.986     & 0.984  & 0.985 & 4031 & 3229 \\
+CNN      & 0.991    & 0.985     & 0.990  & 0.987 & 2137 & 3860 \\
+CNN L    & 0.991    & 0.979     & 0.990  & 0.985 & \textbf{2117} & 3750 \\
+CNN P    & \textbf{0.993}    & \textbf{0.990}& \textbf{0.991}  & \textbf{0.990} & 2432 & \textbf{2114} \\
+CNN N    & 0.991    & 0.987     & 0.988  & 0.988 & 2756 & 3312 \\
+CNN L N  & 0.992    & 0.988     & 0.989  & 0.988 & 2500 & 3567 \\
+LSTM     & 0.741    & 0.184     & 0.500  & 0.269 & 169094 & 0 \\ \hline
 \end{tabular}
-\caption{Scores over the test dataset. N = normalized, L = Lower, P = no position embedding.}
+\caption{Scores over the test dataset. \\\hspace{\textwidth}For models: N = normalized, L = Lower, P = no position embedding. \\\hspace{\textwidth}In headers, FN = False Negative, FP = False Positive}
 \label{tab:scores}
@@
 
-While all model using CNN shows improvement over the baseline, the model definitely does not outperform it by a huge margin (\textless 0.02 FScore).
-And for a reason : the baseline already performs nearly perfectly on the test corpus.
+While all models using CNN show improvement over the baseline, the models definitely do not significantly outperform (\textless 0.02 FScore).
+There is a reason for this: the baseline already performs nearly perfectly on the test corpus.
 The dictionary attack using n-grams did actually perform well.
-As a result, we wanted to compare how both models would perform on a secondary test corpus composed by texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same inputs.
-As a result, we created a new corpus with 4 texts and 742 samples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
-No noise characters and no random keeping of words were applied.
+Therefore, an additional evaluation method was constructed.
+The baseline and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same source corpora and as such the same writing, the same vocabulary.
+This new corpus is composed by 4 texts and counts 742 examples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
+Neither noise characters nor random keeping of words were applied.
 
@@
 
-The following inputs has been tagged with the CNN P model.
-Batch are constructed around the regular expression \texttt{\\W} with package \texttt{regex}.
+The following inputs have been tagged with the CNN P model.
+Batches are constructed around the regular expression \texttt{\\W} with package \texttt{regex}.
 This explains why inputs such as \texttt{".i."} are automatically tagged as \texttt{" .
@@
 "} by the tool.
-The input was stripped of its spaces before tagging, we only show the ground truth by commodity.
+The input was stripped of its spaces before tagging, only the ground truth is shown for readability.
 
@@
 
-\subsection{Discussion}
+\subsection{Evaluation on Latin data}
 
-We believe that, aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated from a text point of view and as a NLP task.
-Word segmentation for some text can be even difficult for humanist, and as such, we believe that post-processing of OCR through tools like Boudams can be a better way to achieve data-mining of the dataset.
+For the following evaluations, the same process was deployed: CNN without Position was evaluated against the baseline on both a test set composed by the same texts that the training text, and an unknown corpora composed by unseen texts.
 
-We were surprised by the negligible effects of the different normalization methods (lower-casing; ASCII reduction; both).
-The presence of certain MUFI characters might provide enough information about segmentation and be in enough numbers for them not to impact the network weights.
-
-While the baseline surprised us by performing this well on the test corpus, it definitely performed less well than the CNN on a completely unknown corpus: in this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical ngram approaches.
-In light of the high accuracy of the CNN model, we believe the model should perform the same way independently from the language in Medieval Western Europe, and results in annexes on Latin confirms this for at least 2 other corpora.
-
-\subsection{Conclusion}
-
-Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more important data mining of OCRed manuscript.
-In aftermath, we wonder if the importance of normalization and lowering should be higher depending on the size of the corpora and its content.
-
-\subsection{Acknowledgements}
-
-Boudams has been made possible by two open-source repositories from which I learned and copied bits of implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}.
-This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.
-
-\bibliographystyle{plainnat}
-\bibliography{article}
-
-\appendix\footnotesize
-
-\section{Annex 1 : Confusion of CNN without position embeddings}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{confusion.png}
-  \caption{Confusion matrix of the CNN model without position embedding}
-  \label{fig:confusion_matrix}
-\end{figure}
-
-\section{Annex 2 : Scores on Latin Prose and Poetic Corpora}
+\subsubsection{Latin Prose and Poetic Corpora}
 
@@
 
-The Latin data is much more noisy than the Old French, as it was less curated than the digital edition we found for Old French.
+The Latin data is much more noisy than the Old French, as it was less curated than the digital edition provided for Old French.
 They are part of the Perseus corpus \citet{perseus} and were cut into passages in the context of my thesis.
-The training, evaluation and test corpora are built upon prose work from Cicero and Suetonius.
+The training, evaluation and test corpora are built upon prose works from Cicero and Suetonius.
 The unknown corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc.
@@
 
-\section{Annex 3: Scores on Medieval Latin corpora}
+\subsubsection{Medieval Latin corpora}
 
@@
 \item Number of classes in unknown examples: 472,655 WC; 113,004 WB
-\end{itemize}
+\end{itemize}p
 
@@
 
-\section{Annex 4: Scores on Latin epigraphic corpora}
+\subsubsection{Latin epigraphic corpora}
 
@@
 The epigraphic Latin corpora is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the corpus of unknown is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
-We also decided to evaluate both the baseline and the model on uppercased data, as it would normally be the state the text would be found in.
+Both the baseline and the model were evaluated on uppercase data, as it would normally be the state the text would be found in.
 Each of the corpora presents a high number of unresolved abbreviations (\textit{ie.} one letter words).
@@
 
-While CNN is influenced by
 \textbf{Statistics:}
@@
 
+\subsection{Discussion}
+
+Aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated as a NLP task.
+Word segmentation for some text can be even difficult for humanist, as shown by the manuscript sample, and as such, it seems that the post-processing of OCR through tools like this one can be a better way to achieve data-mining of raw datasets.
+
+The negligible effects of the different normalization methods (lower-casing; ASCII reduction; both) were surprising.
+The presence of certain MUFI characters might provide enough information about segmentation and be of sufficient quantity for them not to impact the network weights.
+
+While the baseline performed unexpectedly well on the test corpus, the CNN model definitely performed better on a completely unknown corpus.
+In this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical n-gram approaches.
+In light of the high accuracy of the CNN model, the model should perform the same way independently of the language in Medieval Western Europe,.
+
+\subsection{Conclusion}
+
+Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more thorough data mining of OCRed manuscript.
+Given the results, studying the importance of normalization and lowering should probably be a further step, as it might be of high influence in smaller corpora.
+
+\subsection{Acknowledgements}
+
+Boudams has been made possible by two open-source repositories from which I learned and copied bits of implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}.
+This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.
+
+\bibliographystyle{plainnat}
+\bibliography{article}
+
+\appendix\footnotesize
+
+\section{Annex 1 : Confusion of CNN without position embeddings}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{confusion.png}
+  \caption{Confusion matrix of the CNN model without position embedding}
+  \label{fig:confusion_matrix}
+\end{figure}
+
 \end{document}

### 2019-06-07 baselines/ws.py

@@ -5,7 +5,7 @@ import json
 
 import wordsegment as ws
 
-corpus = "latin_epigraphy_uppercase"
+corpus = "fro"
 json_cache = os.path.join(os.path.dirname(__file__), corpus+"_" + "ws_cache.json")
 
 
@@ -90,7 +90,7 @@ with open(os.path.join(os.path.dirname(__file__), "..", "datasets", corpus, "tes
         padding = max(len(truth), padding)
 
 # If compute is True, recompute the data without using the cache
-compute = True
+compute = False
 if compute:
     results = list([
         res

### 2019-06-08 boudams/model/lstm.py
Fixed LSTM model

@@ -24,7 +24,6 @@ class Encoder(nn.Module):
     def forward(self, src, src_len):
 
         # src = [src sent len, batch size]
-
         embedded = self.dropout(self.embedding(src))
 
         # embedded = [src sent len, batch size, emb dim]

### 2019-06-08 article/article.tex
Article update

@@
         \addplot table[x=Epoch,y=CNN L N,col sep=comma] {accuracies.csv}; 
-        \legend{CNN, CNN L, CNN P, CNN N, CNN L N}
+        \addplot table[x=Epoch,y=LSTM,col sep=comma] {accuracies.csv}; 
+        \legend{CNN, CNN L, CNN P, CNN N, CNN L N, LSTM}
       \end{axis}
@@
 CNN L N  & 0.992    & 0.988     & 0.989  & 0.988 & 2500 & 3567 \\
-LSTM     & 0.741    & 0.184     & 0.500  & 0.269 & 169094 & 0 \\ \hline
+LSTM     & 0.939    & 0.637     & 0.918  & 0.720 & 21174 & 18662 \\ \hline
 \end{tabular}

### 2019-06-08 article/code_excerpt.py

@@ -1,9 +1,0 @@
-import mufidecode
-import unidecode
-"sot la gnͣt abstinence dess eintes uirges ele  ꝑla"
-mufidecode.mufidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla")
-# ' sot la gnat abstinence dess eintes uirges ele  pla'
-mufidecode.mufidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla", join=False
-# (' ',  's',  'o',  't',  ' ',  'l',  'a',  ' ',  'g',  'n',  'a',  't',  ' ',  'a',  'b',  's',  't',  'i',  'n',  'e',  'n',  'c',  'e',  ' ',  'd',  'e',  's',  's',  ' ',  'e',  'i',  'n',  't',  'e',  's',  ' ',  'u',  'i',  'r',  'g',  'e',  's',  ' ',  'e',  'l',  'e',  ' ',  ' ',  'p',  'l',  'a')
-unidecode.unidecode(" sot la gnͣt abstinence dess eintes uirges ele  ꝑla")
-# ' sot la gnat abstinence dess eintes uirges ele  la'

### 2020-03-30 article/article.tex
Update article

@@
 
-\title{Evaluating Deep Learning Methods for Tokenization of Space-less texts in Old French and Latin}
+\title{Evaluating Deep Learning Methods for Word Segmentation of Scripta Continua Texts in Old French and Latin}
 \author[1]{Thibault Clérice}
@@
 
-It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese for which an already impressive amount of work has been.
-Chinese word segmentation has lately been driven by deep learning methods, specifically ones based on \textit{sequence to sequence translations}: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses BiDirectional GRU and CRF.
+It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese for which an already impressive amount of work has been done.
+Chinese word segmentation has lately been driven by deep learning methods: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses bi-directional GRU and CRF.
 \footnote{\citet{huang2019realistic} actually gave us the denomination used here: word boundary (WB) and word content (WC).}
@@
 
-The model is based on traditional text input encoding where each character is transcoded to an index.
+The model is based on traditional text input encoding where each character is represented as an index.
 Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content (\textit{cf.} Table \ref{lst:input_output_example}.
@@
 Aside from normalizations of the input and output, three different model structures were tested.
-Every model is composed of one encoder, as described below, and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary).
-For final scores, SOS, EOS and PAD were ignored.
+Every model is composed of one encoder, as described below, and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary) \footnote{For final scores, SOS, EOS and PAD were ignored.}.
 
@@
 \begin{itemize}
-    \item 193,734 training examples;
+    \item 193,734 training examples (group of words);
     \item 23,581 validation examples;
@@
 To compare the results, the \texttt{wordsegment} package \citet{WordSegment} was used as a baseline.
+For this purpose, UDPipe (\citet{udpipe:2017}) was evaluated but scores were lower than this baseline: our LSTM and GRU implementations show however the same difficulties while sharing the same apparent architecture (\footnote{An issue regarding parameters or implementations is not to be excluded.}.
 
@@
         \addplot table[x=Epoch,y=LSTM,col sep=comma] {accuracies.csv}; 
-        \legend{CNN, CNN L, CNN P, CNN N, CNN L N, LSTM}
+        \addplot table[x=Epoch,y=GRU,col sep=comma] {accuracies.csv}; 
+        \legend{CNN, CNN L, CNN P, CNN N, CNN L N, LSTM, GRU}
       \end{axis}
@@
 CNN L N  & 0.992    & 0.988     & 0.989  & 0.988 & 2500 & 3567 \\
-LSTM     & 0.939    & 0.637     & 0.918  & 0.720 & 21174 & 18662 \\ \hline
+LSTM     & 0.939    & 0.637     & 0.918  & 0.720 & 21174 & 18662 \\
+GRU     & 0.933    & 0.645     & 0.645  & 0.910 & 23706 & 19427 \\ \hline
 \end{tabular}
@@
 Therefore, an additional evaluation method was constructed.
-The baseline and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts that were not used in training: indeed, the training, dev and test corpus share the same texts while not sharing the same source corpora and as such the same writing, the same vocabulary.
+The baseline and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts of a different domain.
 This new corpus is composed by 4 texts and counts 742 examples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
@@
 The following inputs have been tagged with the CNN P model.
-Batches are constructed around the regular expression \texttt{\\W} with package \texttt{regex}.
+Batches are constructed around the regular expression \texttt{\textbackslash{}W} with package \texttt{regex}.
 This explains why inputs such as \texttt{".i."} are automatically tagged as \texttt{" .

### 2020-03-30 article/article.tex
Small update on some proof-reading post-peer-review

@@
 \usepackage{tabularx}
+\usepackage{titlesec}
+
+\setcounter{secnumdepth}{4}
+\titleformat{\paragraph}
+{\normalfont\normalsize}{\theparagraph}{1em}{}
+\titlespacing*{\paragraph}
+{0pt}{1.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+
 \newcolumntype{C}{>{\arraybackslash}X} % centered "X" column
@@
 
-Tokenization of space-less strings is a task that is specifically difficult for computers as compared to "whathumanscando".
-\textit{Scripta continua} is a writing phenomenon in which words are separated by spaces that disappeared around the 8th century (see \citet{zanna1998lecture}).
-Nevertheless, spacing can be somewhat erratic in later centuries writings as \citet{stutzmann} explains (\textit{cf.} Figure \ref{fig:4lines}) and it becomes an issue for OCR, as continuous bag of word is only interesting when those are not glued together.
-In the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages can be a pre-processing step for further research to sustain analyses such as authorship attribution or simply allow full-text search.
+Tokenization of spaceless strings is a task that is specifically difficult for computers as compared to "whathumanscando".
+\textit{Scripta continua} is a western Latin writing phenomenon in which words are separated by spaces.
+It disappeared around the 8th century (see \citet{zanna1998lecture}), but, nevertheless, spacing remained erratic \footnote{From a modern point of view.
+What we call here \textit{scripta continua} ranges from traditional \textit{scripta continua} to variable spacing writing.} in later centuries writing as \citet{stutzmann} explains (\textit{cf.} Figure \ref{fig:4lines}).
+The fluctuation of space width, or simply their presence becomes an issue for OCR.
+Indeed, in the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages is quite often a pre-processing step for further research to sustain analyses such as authorship attribution, corpus linguistics or simply allow full-text search \footnote{We have found no previous study of \textit{scripta continua} and the likes as a natural language processing issue, only as an HTR/OCR one, such as \cite{wahlberg2014spotting} and \cite{8269990}}.
 
@@
 It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese for which an already impressive amount of work has been done.
+Indeed, the dimensionality alone of Chinese character set is different from Latin alphabets \footnote{a range of twenty to few hundred characters if we take into account all diacritic combination in Latin scripts, at least few thousands in Chinese, depending on the period \cite{routledge}}, and the important presence of compounds is definitely an issue for segmentation \footnote{According to \cite{tse2017chinese}, \textit{"about 73.6 \% of modern Chinese words are two-character compound words".}}.
 Chinese word segmentation has lately been driven by deep learning methods: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses bi-directional GRU and CRF.
@@
 
-For evaluation purposes, and to reduce the number of input classes, two options for data transcoding were used: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
+For evaluation purposes, and to reduce the amount of input classes, two options for data transcoding were used: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
 \ref{fig:normalization}).
@@
 
-Aside from normalizations of the input and output, three different model structures were tested.
+Aside from normalization of the input and output, three different model structures were tested.
 Every model is composed of one encoder, as described below, and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary) \footnote{For final scores, SOS, EOS and PAD were ignored.}.
 
-The encoders are the following (configurations in parenthesis):
+The encoders are the following (configurations in parentheses):
 
@@
 
-\subsubsection{Datasets}
+% We evaluated the tool on different languages, namely Old French and Latin from different eras or writing traditions.
+Old French is the only language on which all model configurations were tested, Latin was tested only on the best performing configuration against the baseline.
+
+\subsubsection{Evaluation on Old French Data}
+
+\paragraph{Main Dataset}
 
@@
     \item Number of classes in testing examples: 482,776 WC; 169,094 WB
-    \item Number of classes in unknown examples: 26,393 WC; 10,193 WB
 \end{itemize}
@@
 They are between 2 and 8 words-length.
-In order to recreate the condition of OCR noise, dots were added randomly (0.2) between words.
-In order to augment the dataset, words are randomly (0.1) passed over the next example\footnote{This data augmentation was limited to one word per sample}.
+In order to recreate the condition of OCR noise, full stop characters were added randomly (20\% chance) between words.
+In order to augment the dataset, words are randomly (10\% chance) passed over the next example\footnote{This data augmentation was limited to one word per sample}.
 If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain, independently of the maximum number of words.
@@
 
-\subsubsection{Results}
+\paragraph{Results}
 
@@
 
-\subsubsection{Unknown texts}
+\paragraph{Out-of-domain (OUD) texts}
 
@@
 The baseline and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts of a different domain.
-This new corpus is composed by 4 texts and counts 742 examples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
+This new corpus is composed of 4 texts and counts 742 examples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
 Neither noise characters nor random keeping of words were applied.
+The resulting corpus contains 26,393 WC and 10,193 WB.
 
@@
 \end{tabular}
-\caption{Scores over the unknown dataset. FN = False Negative, FP = False Positive}
+\caption{Scores over the out-of-domain dataset. FN = False Negative, FP = False Positive}
 \label{tab:scores_unknown}
@@
 
-\subsubsection{Example of outputs}
+\paragraph{Example of Outputs}
 
@@
 
-\subsection{Evaluation on Latin data}
+\subsubsection{Evaluation on Latin data}
 
-For the following evaluations, the same process was deployed: CNN without Position was evaluated against the baseline on both a test set composed by the same texts that the training text, and an unknown corpora composed by unseen texts.
+For the following evaluations, the same process was deployed: CNN without Position was evaluated against the baseline on both a test set composed by the same texts that the training text, and an out-of-domain corpus composed by unseen texts.
+Evaluation has been done on three different kind of Latins (edited, classical Latin (1); medieval Latin of charters (2); epigraphic Latin (3)) as they show different level of difficulties: they always present rich morphology, but medieval Latin displays spelling variations while epigraphic Latin displays both spelling variation and a high presence of abbreviations.
 
-\subsubsection{Latin Prose and Poetic Corpora}
+\paragraph{Latin Prose and Poetic Corpora}
 
@@
 CNN P & Test & \textbf{0.992} & \textbf{0.987} & \textbf{0.989} & \textbf{0.988} & \textbf{439} & \textbf{584} \\ \hline
-Baseline & Unknown & 0.933 & 0.897 & 0.890 & 0.893 & 1587 & 1409 \\
-CNN P & Unknown & 0.970 & 0.952 & 0.956 & 0.954 & 600 & 709 \\ \hline
+Baseline & OUD & 0.933 & 0.897 & 0.890 & 0.893 & 1587 & 1409 \\
+CNN P & OUD & 0.970 & 0.952 & 0.956 & 0.954 & 600 & 709 \\ \hline
 \end{tabular}
@@
 
-The Latin data is much more noisy than the Old French, as it was less curated than the digital edition provided for Old French.
+The Latin data is much noisier than the Old French, as it was less curated than the digital edition provided for Old French.
 They are part of the Perseus corpus \citet{perseus} and were cut into passages in the context of my thesis.
 The training, evaluation and test corpora are built upon prose works from Cicero and Suetonius.
-The unknown corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc.
-Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+The out-of-domain corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc.
+Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
 
@@
 \item Number of classes in testing examples: 105,915 WC; 26,404 WB
-\item Number of classes in unknown examples: 35,910 WC; 8,828 WB
+\item Number of classes in OUD examples: 35,910 WC; 8,828 WB
 \end{itemize}
@@
 
-\subsubsection{Medieval Latin corpora}
+\paragraph{Medieval Latin corpora}
 
@@
 CNN P & Test & \textbf{0.997} & \textbf{0.995} & \textbf{0.995} & \textbf{0.995} & \textbf{251} & \textbf{298} \\ \hline
-Baseline & Unknown & 0.929 & 0.900 & 0.865 & 0.881 & 14,382 & 27,019 \\
-CNN P & Unknown & 0.976 & 0.960 & 0.963 & 0.962 & 6509 & 7444\\ \hline
+Baseline & OUD & 0.929 & 0.900 & 0.865 & 0.881 & 14,382 & 27,019 \\
+CNN P & OUD & 0.976 & 0.960 & 0.963 & 0.962 & 6509 & 7444\\ \hline
 \end{tabular}
@@
 
-The medieval Latin corpora is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the unknown corpora is based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
-Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+The medieval Latin corpus is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the out-of-domain corpora is based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
+Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
 The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.
@@
 \item Number of classes in testing examples: 137,465 WC; 34,053 WB
-\item Number of classes in unknown examples: 472,655 WC; 113,004 WB
-\end{itemize}p
+\item Number of classes in OUD examples: 472,655 WC; 113,004 WB
+\end{itemize}
 
@@
 
-\subsubsection{Latin epigraphic corpora}
+\paragraph{Latin epigraphic corpora}
 
@@
 CNN P & Test Uppercase & \textbf{0.979} & \textbf{0.972} & \textbf{0.967} & \textbf{0.969} & \textbf{1715} & \textbf{1275} \\ \hline
-Baseline & Unknown & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
-CNN P & Unknown & 0.953 & 0.939 & 0.926 & 0.932 & 4689 & 3112 \\ \hline
-Baseline & Unknown Uppercase & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
-CNN P & Unknown Uppercase & 0.936 & 0.914 & 0.902 & 0.908 & 6152 & 4464 \\ \hline
+Baseline & OUD & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
+CNN P & OUD & 0.953 & 0.939 & 0.926 & 0.932 & 4689 & 3112 \\ \hline
+Baseline & OUD Uppercase & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
+CNN P & OUD Uppercase & 0.936 & 0.914 & 0.902 & 0.908 & 6152 & 4464 \\ \hline
 \end{tabular}
@@
 
-The epigraphic Latin corpora is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the corpus of unknown is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
+The Latin epigraphic corpus is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the out-of-domain corpus is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
 Both the baseline and the model were evaluated on uppercase data, as it would normally be the state the text would be found in.
 Each of the corpora presents a high number of unresolved abbreviations (\textit{ie.} one letter words).
-Both corpus were generated without noise and word keeping, with a maximum sample size of 150 characters.
+Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
 The data presents some polytonic Greek characters, some sample being only in Greek.
@@
 \item Number of classes in testing examples: 107,963 WC; 31,900 WB
-\item Number of classes in unknown examples: 127,268 WC; 38,055 WB
+\item Number of classes in OUD examples: 127,268 WC; 38,055 WB
 \end{itemize}
@@
 
-While the baseline performed unexpectedly well on the test corpus, the CNN model definitely performed better on a completely unknown corpus.
+While the baseline performed unexpectedly well on the test corpus, the CNN model definitely performed better on a out-of-domain corpus.
 In this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical n-gram approaches.
-In light of the high accuracy of the CNN model, the model should perform the same way independently of the language in Medieval Western Europe,.
+In light of the high accuracy of the CNN model, the model should perform the same way independently of the language in Medieval Western Europe.
 
@@
 
-Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more thorough data mining of OCRed manuscript.
+Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more thorough data mining of OCRed manuscripts.
 Given the results, studying the importance of normalization and lowering should probably be a further step, as it might be of high influence in smaller corpora.
@@
 
-Boudams has been made possible by two open-source repositories from which I learned and copied bits of implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}.
+Boudams has been made possible by two open-source repositories from which I learned and copied bits of the implementation of certain modules and without which none of this paper would have been possible: \citet{enrique_manjavacas_2019_2654987} and \citet{bentrevett}.
 This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.

### 2021-05-12 boudams/dataset/plaintext.py
fix segmentation in convert

@@ -9,7 +9,7 @@ from typing import Iterable, Union
 from boudams.dataset.base import write_sentence, _space
 
 
-_splitter = re.compile("(\W+)")
+_splitter = re.compile(r"([^\p{l}\p{p}]+)")
 
 
 def convert(

### 2021-05-12 boudams/dataset/plaintext.py
even better

@@ -9,8 +9,8 @@ from typing import Iterable, Union
 from boudams.dataset.base import write_sentence, _space
 
 
-_splitter = re.compile(r"([^\p{l}\p{p}]+)")
-
+#_splitter = re.compile(r"([^\p{l}\p{p}]+)")
+_splitter = re.compile(r"(\s+)")
 
 def convert(
         input_path: Union[Iterable[str], str], output_path: str,

### 2021-05-13 article/article.tex

@@
 \author[1]{Thibault Clérice}
-\affil[1]{École nationale des Chartes, France} 
-\affil[2]{Université Lyon 3, France}
+\affil[1]{Centre Jean-Mabillon, École nationale des Chartes, PSL Research University, France} 
+\affil[2]{Hisoma (UMR 5189), Université Lyon 3, Université de Lyon (UDL), France}
 
@@
 
-\abstract{Tokenization of modern and old Western European languages seems to be fairly simple, as it stands on the presence mostly of markers such as spaces and punctuation.
-However, when dealing with old sources like manuscripts written in \textit{scripta continua}, antiquity epigraphy or Middle Age manuscripts, (1) such markers are mostly absent, (2) spelling variation and rich morphology make dictionary based approaches difficult.
+\abstract{Tokenization of modern and old Western European languages seems to be fairly simple, as it relies on the presence of markers such as spaces and punctuation.
+However, when dealing with old sources like manuscripts written in \textit{scripta continua}, ancient epigraphy or Middle Age manuscripts, (1) such markers are mostly absent, (2) spelling variation and rich morphology make dictionary based approaches difficult.
 Applying convolutional encoding to characters followed by linear categorization to word-boundary or in-word-sequence is shown to be effective at tokenizing such inputs.
-Additionally, the software is released with a simple interface for tokenizing a corpus or generating a training set.}
+Additionally, the software created for this article (Boudams) is released with a simple interface for tokenizing a corpus or generating a training set \footnote{The software has ongoing development on https://github.com/ponteineptique/boudams .
+The software at the time of the article is also available (\cite{boudams})}.}
 
@@
 Tokenization of spaceless strings is a task that is specifically difficult for computers as compared to "whathumanscando".
-\textit{Scripta continua} is a western Latin writing phenomenon in which words are separated by spaces.
+\textit{Scripta continua} is a western Latin writing phenomenon in which words are not separated by spaces.
 It disappeared around the 8th century (see \citet{zanna1998lecture}), but, nevertheless, spacing remained erratic \footnote{From a modern point of view.
@@
 The fluctuation of space width, or simply their presence becomes an issue for OCR.
-Indeed, in the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages is quite often a pre-processing step for further research to sustain analyses such as authorship attribution, corpus linguistics or simply allow full-text search \footnote{We have found no previous study of \textit{scripta continua} and the likes as a natural language processing issue, only as an HTR/OCR one, such as \cite{wahlberg2014spotting} and \cite{8269990}}.
+Indeed, in the context of text mining of HTR or OCR output, lemmatization and tokenization of medieval western languages is quite often a pre-processing step for further research to sustain analyses such as authorship attribution, corpus linguistics or simply to allow full-text search \footnote{We have found no previous study of \textit{scripta continua} and the likes as a natural language processing issue, only as an HTR/OCR one, such as \cite{wahlberg2014spotting} and \cite{8269990}}.
 
@@
 
-It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese for which an already impressive amount of work has been done.
-Indeed, the dimensionality alone of Chinese character set is different from Latin alphabets \footnote{a range of twenty to few hundred characters if we take into account all diacritic combination in Latin scripts, at least few thousands in Chinese, depending on the period \cite{routledge}}, and the important presence of compounds is definitely an issue for segmentation \footnote{According to \cite{tse2017chinese}, \textit{"about 73.6 \% of modern Chinese words are two-character compound words".}}.
+It must be stressed in this study that the difficulty inherent to segmentation is different for \textit{scripta continua} than the one for languages such as Chinese, for which an already impressive amount of work has been done.
+Indeed, the dimensionality alone of the Chinese character set is different from Latin alphabets \footnote{a range of twenty to a few hundred characters if we take into account all diacritic combination in Latin scripts, at least few thousands in Chinese, depending on the period \cite{routledge}}, and the important presence of compound words  is definitely an issue for segmentation \footnote{According to \cite{tse2017chinese}, \textit{"about 73.6 \% of modern Chinese words are two-character compound words".}}.
 Chinese word segmentation has lately been driven by deep learning methods: \citet{chen2015long} defines a process based on LSTM model, while \citet{yu2019learning} uses bi-directional GRU and CRF.
@@
 Indeed, while the issue with Chinese seems to lie in the decomposition of relatively fixed characters, Old French or Medieval Latin present heavy variation of spelling.
-In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spellings of the word "\textit{cheval}" (horse in Old and Modern French) whose apparition counts span from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}.
+In \citet{camps_pandora}, Camps notes, in the same corpus, the existence of not less than 29 spellings of the word "\textit{cheval}" (horse in Old and Modern French) whose occurrence counts range from 3907 to 1\footnote{These are \textit{cheval}, \textit{chevaus}, \textit{cheual}, \textit{ceval}, \textit{chevals}, \textit{cevaus}, \textit{chival}, \textit{ceual}, \textit{cheuaus}, \textit{cevals}, \textit{chaval}, \textit{chivaus}, \textit{chiual}, \textit{chevas}, \textit{cheuals}, \textit{chiuaus}, \textit{ceuaus}, \textit{chevaul}, \textit{chiuau}, \textit{chivals}, \textit{chevau}, \textit{kevaus}, \textit{chavaus}, \textit{cheuas}, \textit{keval}, \textit{cheua}, \textit{cheuau}, \textit{cheva}, \textit{chiuals}}.
 This  makes a dictionary-based approach rather difficult as it would rely on a high number of different spellings, making the computation highly complex.
@@
 
-The model is based on traditional text input encoding where each character is represented as an index.
+The model used in this study is based on traditional text input encoding where each character is represented as an index.
 Output of the model is a mask that needs to be applied to the input: in the mask, characters are classified either as word boundary or word content (\textit{cf.} Table \ref{lst:input_output_example}.
@@
 
-For evaluation purposes, and to reduce the amount of input classes, two options for data transcoding were used: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
+For evaluation purposes, and to reduce the number of input classes, two options for data transcoding were used: a lower-case normalization and a "reduction to the ASCII character set" feature (fr.
 \ref{fig:normalization}).
-On this point, a lot of issues were encountered with transliteration of medieval paelographic characters that were part of the original datasets, as they are poorly interpreted by the \texttt{unidecode} python package.
-Indeed, \texttt{unidecode} will simply remove characters it does not understand.
-A derivative package named \texttt{mufidecode} was built for this reason(\citet{thibault_clerice_2019_3237731}): it takes precedent over \texttt{unidecode} equivalency tables when the data is known of the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).
+On this latter point, several issues were encountered with the transliteration of medieval paelographic characters that were part of the original datasets, as they are poorly interpreted by the \texttt{unidecode} python package.
+Indeed, \texttt{unidecode} simply removed characters it did not understand.
+A derivative package named \texttt{mufidecode} was built for this reason(\citet{thibault_clerice_2019_3237731}): it takes precedent over \texttt{unidecode} equivalency tables when the character is a known entity in the Medieval Unicode Font Initiative (MUFI, \citet{mufi}).
 
@@
 Aside from normalization of the input and output, three different model structures were tested.
-Every model is composed of one encoder, as described below, and one Linear Classifier which classifies into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary) \footnote{For final scores, SOS, EOS and PAD were ignored.}.
+Every model was composed of one encoder, as described below, and one Linear Classifier which classified characters into 5 classes : Start of Sentence (= SOS), End of Sentence (= EOS), Padding (= PAD), Masked Token (= Word Content), Space (= Word Boundary) \footnote{For final scores, SOS, EOS and PAD were ignored.}.
 
-The encoders are the following (configurations in parentheses):
+The following encoders were used (configurations in parentheses):
 
@@
 
-The dataset is composed of transcriptions (from different projects) of manuscripts with unresolved abbreviation.
+The dataset is composed of transcriptions (from different projects) of manuscripts with unresolved abbreviations.
 The \textbf{Old French} is based on \citet{8269990}, \citet{pinche:hal-01628533}, \citet{jean_baptiste_camps_2019_2630574}, \citet{bfmmss}, and \citet{tnah_transcription}.
@@
 Examples were generated automatically.
-They are between 2 and 8 words-length.
+They are between 2 and 8 words in length.
 In order to recreate the condition of OCR noise, full stop characters were added randomly (20\% chance) between words.
-In order to augment the dataset, words are randomly (10\% chance) passed over the next example\footnote{This data augmentation was limited to one word per sample}.
+In order to augment the dataset, words were randomly (10\% chance) copied from one sample to the next \footnote{This data augmentation was limited to one word per sample.
+\textit{e.g.} If the phrase "I have lived here for a long time" were broken into "I have lived here" and "for a long time", the word "here" might be copied to the second sample, thus producing "I have lived here" and "here for a long time."}.
 If a minimum size of 7 characters was not met in the input sample, another word would be added to the chain, independently of the maximum number of words.
-The example however should not go beyond 100 characters.
-The results corpora should be varied in sizes as shown by Figure \ref{fig:word_sizes}.
-The corpora contains 193 different characters when not normalized, in which some MUFI characters appears few hundred times (\textit{cf.} Table \ref{tab:mufi_examples}).
+The examples, however, could not contain more than 100 characters.
+The word lengths in the results corpora were expected to vary as shown by Figure \ref{fig:word_sizes}.
+The corpora contained 193 different characters when not normalized, in which certain MUFI characters appeared a few hundred times (\textit{cf.} Table \ref{tab:mufi_examples}).
 
@@
 
-The training parameters were 0.00005 in learning rate for each CNN model, 0.001 for the LSTM model, and batch sizes of 64.
+The training parameters were 0.00005 in learning rate for each CNN model (\cite{lecun1998gradient}), 0.001 for the LSTM model (\cite{hochreiter1997long}), and batch sizes of 64.
 Training reached a plateau fairly quickly for each model (\textit{cf.} \ref{fig:loss}).
-Each model except LSTM reached a really low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores}).
+Each model except LSTM achieved a relatively low loss and a high accuracy on the test set (\textit{cf.} \ref{tab:scores}).
 To compare the results, the \texttt{wordsegment} package \citet{WordSegment} was used as a baseline.
@@
           xlabel=Epoch, % Set the labels
-          ylabel=Accuracy,
+          ylabel=Loss,
           legend style={at={(0.5,-0.2)},anchor=north},
@@
         ]
-        \addplot table[x=Epoch,y=CNN 1,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN L,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN w/o P,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN N,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=CNN L N,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=LSTM,col sep=comma] {accuracies.csv}; 
-        \addplot table[x=Epoch,y=GRU,col sep=comma] {accuracies.csv}; 
+        \addplot table[x=Epoch,y=CNN 1,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=CNN L,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=CNN w/o P,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=CNN N,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=CNN L N,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=LSTM,col sep=comma] {loss.csv}; 
+        \addplot table[x=Epoch,y=GRU,col sep=comma] {loss.csv}; 
         \legend{CNN, CNN L, CNN P, CNN N, CNN L N, LSTM, GRU}
@@
     \end{tikzpicture}
-    \caption{Training Loss (Cross-entropy) until plateau was reached. N = normalized, L = Lower, P = no position embedding. LSTM was removed as it did not go below 0.65}
+    \caption{Training Loss (Cross-entropy) until plateau was reached. N = normalized, L = Lower, P = no position embedding.}
   \label{fig:loss}
@@
 
-While all models using CNN show improvement over the baseline, the models definitely do not significantly outperform (\textless 0.02 FScore).
+While all models using CNN showed improvement over the baseline, none of them significantly outperformed it, with a maximum improvement of 0.005 FScore.
 There is a reason for this: the baseline already performs nearly perfectly on the test corpus.
-The dictionary attack using n-grams did actually perform well.
 Therefore, an additional evaluation method was constructed.
-The baseline and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts of a different domain.
-This new corpus is composed of 4 texts and counts 742 examples : the diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
+The baseline (which is dictionary based) and the best achieving deep-learning model (CNN P) were evaluated on a secondary test corpus composed of texts of a different domain.
+This new corpus is composed of 4 texts and counts 742 examples : a diplomatic edition of the \textit{Graal} (\citet{graal}), a \textit{Passion} and a \textit{Vie de Saint Leger}  (\citet{old_french_corpus}), and a \textit{Vie de Saint Thibaut} (\citet{theobaldus}).
 Neither noise characters nor random keeping of words were applied.
@@
 
-The results here were highly different (\textit{cf.} Table \ref{tab:scores_unknown}): while it appears that the CNN is able to expand its "comprehension" of the language to newer texts, the new words are more difficult to take into account for the baseline \texttt{wordsegment} n-gram approach, resulting in a respective drop to 0.945 and 0.838 FScore.
-WordSegment specifically performed badly with WB false positives : it had 3658 over a corpus containing 10,193 WB token (around 35 \%).
+The results here were significantly different (\textit{cf.} Table \ref{tab:scores_unknown}): while the CNN was able to expand its "comprehension" of the language to newer texts, the baseline \texttt{wordsegment} n-gram approach had difficulty dealing with the new vocabulary.
+This resulted in a drop in the FScore to 0.945 for CNN and 0.838 for the baseline.
+WordSegment performed particularly poorly with WB false negatives : it had 3658 over a corpus containing 10,193 WB token (to put it simply, around 35 \% of the spaces were not identified).
 
@@
  & Accuracy & Precision & Recall & FScore & WB FN & WB FP \\ \hline
-Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & 644 \\
-CNN P & \textbf{0.957} & \textbf{0.948} & \textbf{0.944} & \textbf{0.945} & \textbf{854} & \textbf{723} \\ \hline
+Baseline & 0.882 & 0.893 & 0.808 & 0.838 & 3658 & \textbf{644} \\
+CNN P & \textbf{0.957} & \textbf{0.948} & \textbf{0.944} & \textbf{0.945} & \textbf{854} & 723 \\ \hline
 \end{tabular}
-\caption{Scores over the out-of-domain dataset. FN = False Negative, FP = False Positive}
+\caption{Scores (Macro-average) over the out-of-domain dataset. FN = False Negative, FP = False Positive.}
 \label{tab:scores_unknown}
@@
 \hline
-\textbf{Ground truth} & \textbf{Tokenized output} \\\hline
+\textbf{Ground truth} & \textbf{Tokenized output (CNN P)} \\\hline
 Aies joie et leesce en ton cuer car tu auras une fille qui aura .i.
@@
 
-For the following evaluations, the same process was deployed: CNN without Position was evaluated against the baseline on both a test set composed by the same texts that the training text, and an out-of-domain corpus composed by unseen texts.
-Evaluation has been done on three different kind of Latins (edited, classical Latin (1); medieval Latin of charters (2); epigraphic Latin (3)) as they show different level of difficulties: they always present rich morphology, but medieval Latin displays spelling variations while epigraphic Latin displays both spelling variation and a high presence of abbreviations.
+For the following evaluations, the same process was deployed: CNN without Position was evaluated against the baseline on both a test set composed of excerpts from the texts of the training set, and an out-of-domain corpus composed of unseen texts.
+Evaluation has been done on three different categories of Latin texts (edited, classical Latin (1); medieval Latin of charters (2); epigraphic Latin (3)) as they show different levels of difficulty: they always present rich morphology, but medieval Latin displays spelling variations while epigraphic Latin displays both spelling variation and a high number of abbreviations.
 
@@
 
-The Latin data is much noisier than the Old French, as it was less curated than the digital edition provided for Old French.
-They are part of the Perseus corpus \citet{perseus} and were cut into passages in the context of my thesis.
-The training, evaluation and test corpora are built upon prose works from Cicero and Suetonius.
-The out-of-domain corpus is built upon \textit{Epigrammata} from Martial, from book 1 to book 2, as it should be fairly different in word order, vocabulary, etc.
+The Latin data is much noisier than the Old French, as it was less curated than the digital editions provided for Old French.
+They are part of the Perseus corpus \citet{perseus}.
+The training, evaluation and test corpora contain prose works from Cicero and Suetonius.
+The out-of-domain corpus comes from the \textit{Epigrammata} from Martial, from book 1 to book 2, which should be fairly different from the test corpus in word order, vocabulary, etc.
 Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
@@
 
-The medieval Latin corpus is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the out-of-domain corpora is based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
+The medieval Latin corpus is based on the project Formulae - Litterae - Chartae's open data (\citet{formulae}) for its training, evaluation and test sets; the out-of-domain corpora are based on three texts from the Monumenta Germanica (\citet{germanica}) that are from early to late medieval period (Andreas Agnellus, Manegaldus, Theodoricus de Niem) and are drawn from the Corpus Corporum Project.
 Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
-The data presents some MUFI characters but still look like mostly normalized editions, unlike the Old French data.
+The data presents some MUFI characters but otherwise mostly resembles normalized editions, unlike the Old French data.
 
@@
 Baseline & OUD & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
-CNN P & OUD & 0.953 & 0.939 & 0.926 & 0.932 & 4689 & 3112 \\ \hline
+CNN P & OUD & \textbf{0.953} & \textbf{0.939} & \textbf{0.926} & \textbf{0.932} & \textbf{4689} & \textbf{3112} \\ \hline
 Baseline & OUD Uppercase & 0.879 & 0.834 & 0.817 & 0.825 & 8693 & 11332 \\
-CNN P & OUD Uppercase & 0.936 & 0.914 & 0.902 & 0.908 & 6152 & 4464 \\ \hline
+CNN P & OUD Uppercase & \textbf{0.936} & \textbf{0.914} & \textbf{0.902} & \textbf{0.908} & \textbf{6152} & \textbf{4464} \\ \hline
 \end{tabular}
@@
 
-The Latin epigraphic corpus is based on the Epigraphic Database Heidelberg open data \citet{formulae} for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the out-of-domain corpus is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
-Both the baseline and the model were evaluated on uppercase data, as it would normally be the state the text would be found in.
+The Latin epigraphic corpus is based on the Epigraphic Database Heidelberg open data for its training, evaluation and test sets (HD000001-HD010000 and HD010001-HD020000 from \citet{edh}) while the out-of-domain corpus is drawn from an automatic conversion of the Pompei Inscriptions (\citet{pompei}).
+Both the baseline and the model were evaluated on uppercase data, as the texts would normally be found in this state.
 Each of the corpora presents a high number of unresolved abbreviations (\textit{ie.} one letter words).
 Both corpora were generated without noise and word keeping, with a maximum sample size of 150 characters.
-The data presents some polytonic Greek characters, some sample being only in Greek.
+The data present some polytonic Greek characters, since some sample were only in Greek.
 
@@
 
-Aside from a graphical challenge, word segmentation in OCR from manuscripts can actually be treated as a NLP task.
-Word segmentation for some text can be even difficult for humanist, as shown by the manuscript sample, and as such, it seems that the post-processing of OCR through tools like this one can be a better way to achieve data-mining of raw datasets.
+As opposed to being a graphical challenge, word segmentation in OCR from manuscripts can actually be treated as an NLP task.
+Word segmentation for some texts can even be difficult for humanists, as shown by the manuscript sample, and as such, it seems that the post-processing of OCR or HTR through tools like this one can enhance data-mining of raw datasets.
 
@@
 
-While the baseline performed unexpectedly well on the test corpus, the CNN model definitely performed better on a out-of-domain corpus.
-In this context, the proposed model actually shows its ability to carry over unknown corpora in a better way than classical n-gram approaches.
-In light of the high accuracy of the CNN model, the model should perform the same way independently of the language in Medieval Western Europe.
+While the baseline performed unexpectedly well on the test corpora, the CNN model definitely performed better on the out-of-domain corpora.
+In this context, the proposed model deals better with unknown corpora classical n-gram approaches.
+In light of the high accuracy of the CNN model on the different corpora, the model should perform equally well no matter to which Medieval Western European language it is applied.
 
@@
 Achieving 0.99 accuracy on word segmentation with a corpus as large as 25,000 test samples seems to be the first step for a more thorough data mining of OCRed manuscripts.
-Given the results, studying the importance of normalization and lowering should probably be a further step, as it might be of high influence in smaller corpora.
+Given the results, studying the importance of normalization and lowering should be the next step, as it will probably show greater influence in smaller corpora.
 
@@
 This tool was originally intended for post-processing OCR for the presentation \citet{pinchecampsclerice} at DH2019 in Utrecht.
+The software was developed using PyTorch (\cite{pytorch}), Numpy (\cite{numpy}) and SK-Learn (\cite{pedregosa2011scikit}).
 

### 2021-05-13 boudams/dataset/plaintext.py
Fix segmentation in convert (#18)
* fix segmentation in convert
* even better

@@ -9,9 +9,8 @@ from typing import Iterable, Union
 from boudams.dataset.base import write_sentence, _space
 
 
-_splitter = re.compile("(\W+)")
+_splitter = re.compile(r"(\s+)")
 _apos = re.compile("['’]")
-
 
 def convert(
         input_path: Union[Iterable[str], str], output_path: str,
